### 2018-07-05 doc/manuscript.tex
working on report

\documentclass[11pt,a4paper,draft=false]{report}
\usepackage{amsmath,amsthm,amssymb}
\usepackage{bm,txfonts}       % bm: bold for Greek letters, txfonts -> varmathbb
\usepackage{breqn}              % automatically brake line in equations
\usepackage{cancel}
\usepackage{arydshln}         % matrix dash lines
\usepackage{booktabs}         % tables
\usepackage{color,soul}         % used for \hl{}
\usepackage[inline]{enumitem}   % use to set labels in the enumerate environment
\usepackage{fullpage}
\usepackage[pdftex]{graphicx}
\graphicspath{{./figures/}{./tikz/}}
\DeclareGraphicsExtensions{.pdf,.png,.jpg}
\usepackage{caption}
\captionsetup{labelfont=bf}
\usepackage[caption=false,labelfont=bf]{subfig}
\usepackage{listings}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{url}
\usepackage[pdfpagelabels,pdfusetitle,colorlinks=true,pdfborder={0 0
  0}]{hyperref}
\usepackage{natbib}
\bibliographystyle{IEEEtranN}

% indexing
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% \usepackage{makeidx}
% \makeindex

% make use of \index{word} in text

% acronyms
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage[acronym]{glossaries}

\newacronym{tscmc}{TSCMC}{Task Space Computed Muscle Control}
\newacronym{cmc}{CMC}{Computed Muscle Control}
\newacronym{dofs}{DoFs}{Degrees of Freedom}
\newacronym{dof}{DoF}{Degree of Freedom}
\newacronym{ik}{IK}{Inverse Kinematics}
\newacronym{tsik}{TSIK}{Task Space Inverse Kinematics}
\newacronym{rms}{RMS}{Root Mean Square}
\newacronym{fd}{FD}{Forward Dynamics}
\newacronym{id}{ID}{Inverse Dynamics}
\newacronym{md}{MD}{Mixed Dynamics}
\newacronym{jra}{JRA}{Joint Reaction Analysis}
\newacronym{eoms}{EoMs}{Equations of Motion}
\newacronym{emg}{EMG}{Electromyography}
\newacronym{mocap}{MoCap}{Motion Capture}
\newacronym{imu}{IMU}{Inertial Measurement Unit}
\newacronym{itsc}{ITSC}{Inverse Task Space Controller}
\newacronym{so}{SO}{Static Optimization}
\newacronym{pd}{PD}{Proportional Derivative}
\newacronym{dmc}{DMC}{Direct Marker Control}
\newacronym{dae}{DAE}{Differential Algebraic Equations}
\newacronym{rra}{RRA}{Residual Reduction Algorithm}
\newacronym{ode}{ODE}{Ordinary Differential Equations}
\newacronym{svd}{SVD}{Singular Value Decomposition}
\newacronym{mpp}{MPP}{Moore-Penrose Pseudoinverse}
\newacronym{cns}{CNS}{Central Nervous System}
\newacronym{umh}{UMH}{Uncontrolled Manifold Hypothesis}
\newacronym{ke}{KE}{Kinetic Energy}
\newacronym{pe}{PE}{Potential Energy}
\newacronym{rmse}{RMSE}{Root Mean Square Error}
\newacronym{em}{EOMs}{Extraocular Muscles}
\makeglossaries{}

% custom commands
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\DeclareMathAlphabet\mathbfcal{OMS}{cmsy}{b}{n} % mathcal bf
\newcommand{\ten}[1]{\mathbfcal{#1}}
\newcommand{\mat}[1]{\bm{#1}}
\renewcommand*{\vec}[1]{\bm{#1}}
\newcommand{\pd}[2]{\frac{\partial#1}{\partial#2}}
\newcommand{\subspace}[1]{\varmathbb{#1}}
\newcommand{\rs}[1]{\subspace{R} (#1)}
\newcommand{\cs}[1]{\subspace{C} (#1)}
\newcommand{\ns}[1]{\subspace{N} (#1)}
\newcommand{\fs}[1]{\subspace{F}}
\newcommand{\ms}[1]{\subspace{M}}
\newcommand{\R}[1]{\mathfrak{R}^{#1}}
\newcommand{\inr}[1]{\in\R{#1}}
\newcommand{\TP}[1]{\mat{T}_{#1}}     % tall projection
\newcommand{\FP}[1]{\mat{F}_{#1}}     % fat projection
\newcommand{\NT}[1]{\mat{N}_{\TP{#1}}}
\newcommand{\N}[1]{\mat{N}_{#1}}
\newcommand{\dNT}[1]{\mat{\dot{N}}_{\TP{#1}}}
\newcommand{\NF}[1]{\mat{N}_{\FP{#1}}}
\newcommand{\dNF}[1]{\mat{\dot{N}}_{\FP{#1}}}
\newcommand{\mc}[1]{\mathcal{#1}}

% mathematical proofs
\newtheorem{theorem}{Theorem}[chapter]
\newtheorem{corollary}{Corollary}[chapter]
\newcommand{\corollaryautorefname}{Corollary}
\newtheorem{proposition}{Proposition}[chapter]
\newcommand{\propositionautorefname}{Proposition}
\newtheorem{remark}{Remark}[chapter]
\newcommand{\remarkautorefname}{Remark}
\newtheorem{definition}{Definition}[chapter]
\newcommand{\definitionautorefname}{Definition}
\newcommand{\subfigureautorefname}{Figure} % subfloat
\renewcommand*{\figureautorefname}{Figure}
\renewcommand*{\sectionautorefname}{Section}
\renewcommand*{\subsectionautorefname}{Subsection}
\newcommand{\algorithmautorefname}{Algorithm}
\makeatletter
\renewcommand\thealgorithm{\thechapter.\arabic{algorithm}}
\@addtoreset{algorithm}{chapter}
\makeatother
\def\equationautorefname~#1\null{Equation #1\null}

% \renewcommand{\baselinestretch}{1.5} % line stretch

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{An Open-Source OpenSim Oculomotor Model for Kinematic and Dynamic
  Simulation}

\author{Constantinos Filip, Dimitar Stanev\footnote{Electrical and Computer
    Engineering Department, University of Patras, Greece, Corresponding author:
    stanev@ece.upatras.gr} and Konstantinos Moustakas}

\date{\today}

\begin{document}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
  bla
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Introduction}\label{sec:introduction}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Methods}\label{sec:methods}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Eye Globe}\label{sec:eye-globe}

he orbital plant consists of the globe (eyeball), three pairs of extraocular
muscles, and connective tissues. The size of an emmetropic human adult eye is
approximately 24.2 mm (transverse, horizontal) x 23.7 mm (sagittal, vertical) x
22.0–24.8 mm (axial, anteroposterior) with no significant difference between
sexes and age groups. In the transverse diameter, the eyeball size may vary from
21 mm to 27 mm. Thus, it can be approximated by a solid sphere with 12mm
radius. The eyeball was constructed in Blender, an open-source software 3D
creation Software. We used a spherical mesh with 32 segments and 12 rings, to
construct the vitreous humor (body) as solid sphere and a conical plate to
construct the cornea. The weight of an average human eye is 7.5grams and the
moment of inertia can be calculated similarly as in the case of a spherical
homogenous and isotropic object with radius 12mm ($\mat{I} = 2/5 m r^2$ at the
center of mass).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Muscle Paths and Pulleys}\label{sec:muscle-modeling}

\begin{table}[h]
  \caption{TODO}\label{tab:muscle-path}
  \begin{tabular}{@{}cccccccccc@{}}
    \toprule
    \textbf{Muscle} & \multicolumn{3}{c}{\textbf{Origin}}                                & \multicolumn{3}{c}{\textbf{Pulley}}                                & \multicolumn{3}{c}{\textbf{Insertion}}                                                                      \\ \midrule
                    & \textit{\textbf{Ox}} & \textit{\textbf{Oy}} & \textit{\textbf{Oz}} & \textit{\textbf{Px}} & \textit{\textbf{Py}} & \textit{\textbf{Pz}} & \textit{\textbf{Ix}} & \multicolumn{1}{l}{\textit{\textbf{Iy}}} & \multicolumn{1}{l}{\textit{\textbf{Iz}}} \\ \midrule
    LR              & -0.034               & 0.0006               & -0.013               & -0.0102              & 0.0003               & 0.012                & 0.0065               & \multicolumn{1}{l}{0}                    & \multicolumn{1}{l}{0.0101}               \\
    MR              & -0.030               & 0.0006               & -0.017               & -0.0053              & 0.00014              & -0.0146              & 0.0088               & \multicolumn{1}{l}{0}                    & \multicolumn{1}{l}{-0.0096}              \\
    SR              & -0.0317              & 0.0036               & -0.016               & -0.0092              & 0.012                & -0.002               & 0.0076               & 0.0104                                   & 0                                        \\
    IR              & -0.0317              & -0.0024              & -0.016               & -0.0042              & -0.0128              & -0.0042              & 0.00805              & -0.0102                                  & 0                                        \\
    SO              & 0.0082               & 0.0122               & -0.0152              & -0.030834            & 0.001145             & -0.01644             & 0.0044               & 0.011                                    & 0.0029                                   \\
    IS              & 0.0113               & -0.0154              & -0.0111
                    & -0.00718             & -0.0135              & 0
                                                                                         &
                                                                                           -0.008
                                                                                                                & 0                                        & 0.009 \\
    \bottomrule
  \end{tabular}
\end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Passive Connective Tissues}\label{sec:passive-connective-tissues}

The passive connective tissues of the eyeball apply a restoring force, which
brings the globe back to the central position when the net force from the
\gls{em} is zero. These tissues include all non-muscular suspensory tissues,
such as Tenon’s capsule, the optic nerve, the fat pad and the conjunctiva. The
force-displacement curve of the net elasticity can be represented as

\begin{equation}\label{equ:passive-tissue}
  \vec{f}_t = -k_p \vec{q} - k_c 10^{-3} \vec{q}^3 - k_d * \vec{\dot{q}}
\end{equation}
%
where, $\vec{f}_t$ represents the passive tissue forces,
$k_p= 0.33 \text{g / deg}$, $k_p= 1.56 \text{g / (deg^3)}$ (\hl{TODO: convert to
  Nm / rad}) and
$k_v= 10^{-4} \text{Nm s / rad}$ the constants and $\vec{\dot{q}} \inr{3}$ the
rotational coordinates of the model. These forces serve the eye’s stabilization,
and are modeled using \texttt{OpenSim}'s expression based coordinate force.
% 0.33 * 9.8066500286389 * 10**-3 / (numpy.pi / 180) = 0.18542028707494282
% 1.56 * 9.8066500286389 * 10**-3 / (numpy.pi / 180) ** 3 = 2877.4856893811248

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Model Validation}\label{sec:model-validation}

To meet good fidelity criteria, a model requires to be verified and
validated. In our study, verification in oculomotor models was performed by
comparing the force-length characteristic curves of the modeled \gls{em} to
published data. These data include only the lateral rectus characteristic
curves. However, we can make a safe assumption that the other muscles have
similar properties. More characteristic curves showing the force-length
relationship of the other \gls{em}, as well as the changes in muscle length with
rotation in all directions, are presented in the end of this report. Further
verification can be done by comparing the model’s joint forces produced during
horizontal movement with clinically collected data, but these data are not
available.  Validation was performed by simulating and analyzing synthesized eye
movements while ensuring that Listing Law was obeyed by maintaining zero torsion
in secondary gaze positions, and by testing if the assessed innervations are
sufficient in fixating the eye at a desired position.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Results}\label{sec:results}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \section*{Discussion}\label{sec:discussion}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Conclusion}\label{sec:conclusion}

We manage to construct a complete ocular model that represents the ocular
motility of a normal human eye. The model can be used to drive simulations of
different eye movement systems and be derive some inference about the excitation
and activation patterns. The verification and validation of the model showed
that it matches the data found in literature and it is able to produce
synthetized movements. We even showed that the simulation results produced by
static optimization and forward dynamics in OpenSim are less satisfactory than
the proposed solution where we used a PD controller to minimize the tracking
error between the desired and estimated trajectory of saccadic movements. The
desired kinematics were tracked within a maximum \gls{rmse} of and in horizontal
and vertical saccades respectively. With the application of the controller, a
very rapid instantaneous acceleration and a constant velocity to sustain clear
vision is attained. The produced activation levels were in accordance with the
descriptions found in the bibliography and the highest activation levels were
shown only during the time intervals when the main agonist muscle was activated.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \bibliography{mylibrary}
% \addcontentsline{toc}{chapter}{Bibliography}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:

### 2018-07-05 source_code/src/fixation_controller_plugin/FixationController.cpp

@@ -7,19 +7,19 @@ using namespace OpenSim;
 using namespace SimTK;
 
 /**
-* Implementation of smooth sigmoid function.
-*
-* Parameters
-* ----------
-* t : time to be evaluated
-* t0 : delay
-* A : magnitude
-* B : slope
-*
-* Returns
-* -------
-* (y, y', y'')
-*/
+ * Implementation of smooth sigmoid function.
+ *
+ * Parameters
+ * ----------
+ * t : time to be evaluated
+ * t0 : delay
+ * A : magnitude
+ * B : slope
+ *
+ * Returns
+ * -------
+ * (y, y', y'')
+ */
 Vec3 sigmoid(double t, double t0, double A, double B) {
     return Vec3(A * (tanh(B * (t - t0)) + 1) / 2,
                 A * B * (1 - pow(tanh(B * (t - t0)), 2)) / 2,

### 2018-07-05 source_code/src/fixation_controller_plugin/FixationController.h

@@ -1,11 +1,16 @@
 /**
-* @file FixationController.h
-*
-* \brief A
-*
-* @author Constantinos Filip <filipconstantinos@gmail.com>
-*         Dimitar Stanev     <jimstanev@gmail.com>
-*/
+ * @file FixationController.h
+ *
+ * \brief A fixation controller for the eye model. The controller accepts the
+ * horizontal and vertical target angle and computes the muscle excitation
+ * required to track the desired motion.
+ *
+ * @author Constantinos Filip <filipconstantinos@gmail.com>
+ *         Dimitar Stanev     <jimstanev@gmail.com>
+ *
+ * @see <a href="https://simtk.org/projects/eye">[SimTK Project]</a>, <a
+ * href="XXX">[Publication]</a>
+ */
 #ifndef FIXATION_CONTROLLER_H
 #define FIXATION_CONTROLLER_H
 
@@ -14,9 +19,7 @@
 
 namespace OpenSim {
     /**
-     * \brief A
-     *
-     *
+     * \brief Eye fixation controller.
      */
     class FixationController_API FixationController : public Controller {
         OpenSim_DECLARE_CONCRETE_OBJECT(FixationController, Controller);
@@ -31,11 +34,13 @@ namespace OpenSim {
         OpenSim_DECLARE_PROPERTY(kdT, double, "torsion velocity tracking gain");
         OpenSim_DECLARE_PROPERTY(saccade_onset, double, "saccade onset (s)");
         OpenSim_DECLARE_PROPERTY(saccade_velocity, double, "saccade velocity (in rad/s^2)");
-
+        /** Constructor. */
         FixationController();
+        /** Distractor. */
         ~FixationController() {};
+        /** Required to initialize controller properties. */
         void constructProperties();
-        // Controller::computeControls
+        /** Controller::computeControls. */
         void computeControls(const SimTK::State& s, SimTK::Vector& controls)
             const override;
     };

### 2018-07-05 source_code/src/simulation/Simulation.cpp

@@ -1,10 +1,16 @@
 /**
  * @file Simulation.cpp
  *
- * \brief Simulation
+ * \brief Loads the baseline model, removes the linear tissue forces and adds
+ * and ExpressionBasedCoordinateForce. A custom FixationController is added to
+ * the model and a forward simulation is carried. All results are stored in the
+ * model/ folder.
  *
  * @author Constantinos Filip <filipconstantinos@gmail.com>
  *         Dimitar Stanev     <jimstanev@gmail.com>
+ *
+ * @see <a href="https://simtk.org/projects/eye">[SimTK Project]</a>, <a
+ * href="XXX">[Publication]</a>
  */
 #include <iostream>
 #include <OpenSim/OpenSim.h>
@@ -22,17 +28,19 @@ void addExpressionCoordinateForce(Model* model) {
     // remove other forces first
     model->updForceSet().remove(
         model->updForceSet().getIndex(
-        &model->updForceSet().get("add_adb_tissue"), 0));
+            &model->updForceSet().get("add_adb_tissue"), 0));
 
     model->updForceSet().remove(
         model->updForceSet().getIndex(
-        &model->updForceSet().get("sup_inf_tissue"), 0));
+            &model->updForceSet().get("sup_inf_tissue"), 0));
 
     model->updForceSet().remove(
         model->updForceSet().getIndex(
-        &model->updForceSet().get("inc_exc_tissue"), 0));
+            &model->updForceSet().get("inc_exc_tissue"), 0));
 
-    // add expression based coordinate forces (don't use spaces in the expression)
+    // Add expression based coordinate forces (don't use spaces in the
+    // expression). Linux users may experience problem with Lepton (v3.3), most
+    // recent versions may not have this problem.
     auto add_adb_tissue =
         new ExpressionBasedCoordinateForce("r_eye_add_abd",
                                            "-0.002225*q-0.0105*0.0001*q^3-0.001*qdot");
@@ -79,12 +87,12 @@ void simulateModel() {
     auto& state = model.initSystem();
     model.equilibrateMuscles(state);
     /*model.updCoordinateSet().get("r_eye_add_abd").setValue(state, 0.1);
-    model.updCoordinateSet().get("r_eye_sup_inf").setValue(state, 0.1);*/
+      model.updCoordinateSet().get("r_eye_sup_inf").setValue(state, 0.1);*/
 
     // Visualize model
[…]

### 2018-07-05 doc/manuscript.tex
filip version

@@
 \usepackage{listings}
-\usepackage{algorithm}
-\usepackage{algorithmic}
+% \usepackage{algorithm}
+% \usepackage{algorithmic}
 \usepackage{url}
@@
 \makeatletter
-\renewcommand\thealgorithm{\thechapter.\arabic{algorithm}}
+% \renewcommand\thealgorithm{\thechapter.\arabic{algorithm}}
 \@addtoreset{algorithm}{chapter}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection*{Eye Globe}\label{sec:eye-globe}
+\subsection*{Eye Modeling}\label{sec:eye-Modeling}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection*{Muscle Paths and Pulleys}\label{sec:muscle-modeling}
+\subsection*{Muscle Modeling}\label{sec:muscle-modeling}
 
@@
 where, $\vec{f}_t$ represents the passive tissue forces,
-$k_p= 0.33 \text{g / deg}$, $k_p= 1.56 \text{g / (deg^3)}$ (\hl{TODO: convert to
-  Nm / rad}) and
-$k_v= 10^{-4} \text{Nm s / rad}$ the constants and $\vec{\dot{q}} \inr{3}$ the
+$k_p= 0.33 g / deg$ , $k_p= 1.56 g / (deg^3)$  (\hl{TODO: convert to
+	$Nm / rad$}) and
+$k_v= 10^{-4} Nm s / rad$ the constants and $\vec{\dot{q}} \inr{3}$ the
 rotational coordinates of the model. These forces serve the eye’s stabilization,
@@
 % 1.56 * 9.8066500286389 * 10**-3 / (numpy.pi / 180) ** 3 = 2877.4856893811248
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section*{Results}\label{sec:results}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section*{Results}\label{sec:results}
+\subsection*{Fixation Controller}\label{sec:fixation-controller}
 

### 2018-07-05 source_code/src/include/Settings.h

@@ -3,6 +3,6 @@
 
 #include <string>
 
-std::string MODEL_DIR = "/home/mitkof6/dev/upat_eye_model/source_code/../model/";
+std::string MODEL_DIR = "D:/Desktop/upat_eye_model/source_code/../model/";
 
 #endif

### 2018-07-09 doc/manuscript.tex
manuscript

@@
 
-he orbital plant consists of the globe (eyeball), three pairs of extraocular
+The orbital plant consists of the globe (eyeball), three pairs of extraocular
 muscles, and connective tissues. The size of an emmetropic human adult eye is
-approximately 24.2 mm (transverse, horizontal) x 23.7 mm (sagittal, vertical) x
+approximately 24.2 mm (transverse, horizontal), 23.7 mm (sagittal, vertical),
 22.0–24.8 mm (axial, anteroposterior) with no significant difference between
 sexes and age groups. In the transverse diameter, the eyeball size may vary from
-21 mm to 27 mm. Thus, it can be approximated by a solid sphere with 12mm
+21 mm to 27 mm. Thus, it can be approximated by a solid sphere with 12 mm
 radius. The eyeball was constructed in Blender, an open-source software 3D
@@
 construct the vitreous humor (body) as solid sphere and a conical plate to
-construct the cornea. The weight of an average human eye is 7.5grams and the
+construct the cornea. The weight of an average human eye is 7.5 g and the
 moment of inertia can be calculated similarly as in the case of a spherical
-homogenous and isotropic object with radius 12mm ($\mat{I} = 2/5 m r^2$ at the
+homogenous and isotropic object with radius 12 mm ($\mat{I} = 2/5 m r^2$ at the
 center of mass).
@@
     \toprule
-    \textbf{Muscle} & \multicolumn{3}{c}{\textbf{Origin}}                                & \multicolumn{3}{c}{\textbf{Pulley}}                                & \multicolumn{3}{c}{\textbf{Insertion}}                                                                      \\ \midrule
-                    & \textit{\textbf{Ox}} & \textit{\textbf{Oy}} & \textit{\textbf{Oz}} & \textit{\textbf{Px}} & \textit{\textbf{Py}} & \textit{\textbf{Pz}} & \textit{\textbf{Ix}} & \multicolumn{1}{l}{\textit{\textbf{Iy}}} & \multicolumn{1}{l}{\textit{\textbf{Iz}}} \\ \midrule
-    LR              & -0.034               & 0.0006               & -0.013               & -0.0102              & 0.0003               & 0.012                & 0.0065               & \multicolumn{1}{l}{0}                    & \multicolumn{1}{l}{0.0101}               \\
-    MR              & -0.030               & 0.0006               & -0.017               & -0.0053              & 0.00014              & -0.0146              & 0.0088               & \multicolumn{1}{l}{0}                    & \multicolumn{1}{l}{-0.0096}              \\
+    \textbf{Muscle}
+    & \multicolumn{3}{c}{\textbf{Origin}}
+    & \multicolumn{3}{c}{\textbf{Pulley}}
+    & \multicolumn{3}{c}{\textbf{Insertion}}
+    \\ \midrule
+    & \textit{\textbf{Ox}} & \textit{\textbf{Oy}} & \textit{\textbf{Oz}}
+    & \textit{\textbf{Px}} & \textit{\textbf{Py}} & \textit{\textbf{Pz}}
+    & \textit{\textbf{Ix}} & \textit{\textbf{Iy}} & \textit{\textbf{Iz}}
+    \\ \midrule
+    LR & -0.034 & 0.0006 & -0.013 & -0.0102 & 0.0003 & 0.012 & 0.0065 & 0 & 0.0101 \\
+    MR & -0.030 & 0.0006 & -0.017 & -0.0053 & 0.00014 & -0.0146 & 0.0088 & 0 & -0.0096 \\
     SR & -0.0317 & 0.0036 & -0.016 & -0.0092 & 0.012 & -0.002 & 0.0076 & 0.0104 & 0 \\
@@
     SO & 0.0082 & 0.0122 & -0.0152 & -0.030834 & 0.001145 & -0.01644 & 0.0044 & 0.011 & 0.0029 \\
-    IS              & 0.0113               & -0.0154              & -0.0111
-                    & -0.00718             & -0.0135              & 0
-                                                                                         &
-                                                                                           -0.008
-                                                                                                                & 0                                        & 0.009 \\
+    IS & 0.0113 & -0.0154 & -0.0111 & -0.00718 & -0.0135 & 0 & -0.008 & 0 & 0.009 \\
     \bottomrule
@@
 % 
-where, $\vec{f}_t$ represents the passive tissue forces,
-$k_p= 0.33 g / deg$ , $k_p= 1.56 g / (deg^3)$  (\hl{TODO: convert to
-	$Nm / rad$}) and
-$k_v= 10^{-4} Nm s / rad$ the constants and $\vec{\dot{q}} \inr{3}$ the
-rotational coordinates of the model. These forces serve the eye’s stabilization,
-and are modeled using \texttt{OpenSim}'s expression based coordinate force.
+where, $\vec{f}_t$ represents the passive tissue forces, $k_p= 0.33$ g / deg,
+$k_p= 1.56$ g / (deg^3) (\hl{TODO:convert to $Nm / rad$}) and $k_v= 10^{-4}$ Nm
+s / rad the constants and $\vec{\dot{q}} \inr{3}$ the rotational coordinates of
+the model. These forces serve the eye’s stabilization, and are modeled using
+\texttt{OpenSim}'s expression based coordinate force.
 % 0.33 * 9.8066500286389 * 10**-3 / (numpy.pi / 180) = 0.18542028707494282

### 2018-07-09 source_code/src/simulation/Simulation.cpp
correct passive force parameters

@@ -28,34 +28,34 @@ void addExpressionCoordinateForce(Model* model) {
     // remove other forces first
     model->updForceSet().remove(
         model->updForceSet().getIndex(
-            &model->updForceSet().get("add_adb_tissue"), 0));
+        &model->updForceSet().get("add_adb_tissue"), 0));
 
     model->updForceSet().remove(
         model->updForceSet().getIndex(
-            &model->updForceSet().get("sup_inf_tissue"), 0));
+        &model->updForceSet().get("sup_inf_tissue"), 0));
 
     model->updForceSet().remove(
         model->updForceSet().getIndex(
-            &model->updForceSet().get("inc_exc_tissue"), 0));
+        &model->updForceSet().get("inc_exc_tissue"), 0));
 
     // Add expression based coordinate forces (don't use spaces in the
     // expression). Linux users may experience problem with Lepton (v3.3), most
     // recent versions may not have this problem.
     auto add_adb_tissue =
         new ExpressionBasedCoordinateForce("r_eye_add_abd",
-                                           "-0.002225*q-0.0105*0.0001*q^3-0.001*qdot");
+                                           "-0.002225*q-34.5297*0.0001*q^3-0.002*qdot");
     add_adb_tissue->setName("add_adb_tissue");
     model->addForce(add_adb_tissue);
 
     auto sup_inf_tissue =
         new ExpressionBasedCoordinateForce("r_eye_sup_inf",
-                                           "-0.002225*q-0.0105*0.0001*q^3-0.001*qdot");
+                                           "-0.002225*q-34.5297*0.0001*q^3-0.002*qdot");
     sup_inf_tissue->setName("sup_inf_tissue");
     model->addForce(sup_inf_tissue);
 
     auto inc_exc_tissue =
         new ExpressionBasedCoordinateForce("r_eye_inc_exc",
-                                           "-0.002225*q-0.0105*0.0001*q^3-0.001*qdot");
+                                           "-0.002225*q-34.5297*0.0001*q^3-0.002*qdot");
     inc_exc_tissue->setName("inc_exc_tissue");
     model->addForce(inc_exc_tissue);
 }
@@ -89,12 +89,12 @@ void simulateModel() {
     /*model.updCoordinateSet().get("r_eye_add_abd").setValue(state, 0.1);
       model.updCoordinateSet().get("r_eye_sup_inf").setValue(state, 0.1);*/
 
-    // Visualize model
[…]

### 2018-07-09 doc/manuscript.tex
working on manuscript

@@
 
-\newacronym{tscmc}{TSCMC}{Task Space Computed Muscle Control}
-\newacronym{cmc}{CMC}{Computed Muscle Control}
 \newacronym{dofs}{DoFs}{Degrees of Freedom}
 \newacronym{dof}{DoF}{Degree of Freedom}
-\newacronym{ik}{IK}{Inverse Kinematics}
-\newacronym{tsik}{TSIK}{Task Space Inverse Kinematics}
-\newacronym{rms}{RMS}{Root Mean Square}
-\newacronym{fd}{FD}{Forward Dynamics}
-\newacronym{id}{ID}{Inverse Dynamics}
-\newacronym{md}{MD}{Mixed Dynamics}
-\newacronym{jra}{JRA}{Joint Reaction Analysis}
 \newacronym{eoms}{EoMs}{Equations of Motion}
-\newacronym{emg}{EMG}{Electromyography}
-\newacronym{mocap}{MoCap}{Motion Capture}
-\newacronym{imu}{IMU}{Inertial Measurement Unit}
-\newacronym{itsc}{ITSC}{Inverse Task Space Controller}
-\newacronym{so}{SO}{Static Optimization}
+% \newacronym{so}{SO}{Static Optimization}
 \newacronym{pd}{PD}{Proportional Derivative}
-\newacronym{dmc}{DMC}{Direct Marker Control}
-\newacronym{dae}{DAE}{Differential Algebraic Equations}
-\newacronym{rra}{RRA}{Residual Reduction Algorithm}
-\newacronym{ode}{ODE}{Ordinary Differential Equations}
-\newacronym{svd}{SVD}{Singular Value Decomposition}
-\newacronym{mpp}{MPP}{Moore-Penrose Pseudoinverse}
 \newacronym{cns}{CNS}{Central Nervous System}
-\newacronym{umh}{UMH}{Uncontrolled Manifold Hypothesis}
-\newacronym{ke}{KE}{Kinetic Energy}
-\newacronym{pe}{PE}{Potential Energy}
+\newacronym{lr}{LR}{Lateral Rectus}
+\newacronym{sr}{SR}{Superior Rectus}
+\newacronym{ir}{IR}{Inferior Rectus}
+\newacronym{mr}{MR}{Medial Rectus}
+\newacronym{so}{SO}{Superior Oblique}
+\newacronym{io}{IO}{Inferior Oblique}
 \newacronym{rmse}{RMSE}{Root Mean Square Error}
 \newacronym{em}{EOMs}{Extraocular Muscles}
+\newacronym{fl}{F-L}{Force-Length}
+\newacronym{fv}{F-V}{Force-Velocity}
 \makeglossaries{}
@@
 
+% in order to add newline in table cells
+\usepackage{makecell}
+\renewcommand\theadalign{bc}
+\renewcommand\theadfont{\bfseries}
+\renewcommand\theadgape{\Gape[4pt]}
+\renewcommand\cellgape{\Gape[4pt]}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
@@
 
+Rapid and accurate eye movements are crucial for coordinated direction of
+gaze. Studying human eye movement has significant implications for improving our
+understanding of the oculomotor system and treating visuomotor disorders. Over
+the years, biomechanic simulation has provided an analysis tool of different
+human movements, especially gait. This analysis can be extended to visual tasks
+by analyzing the mechanisms of ocular motility and providing a realistic ocular
+model that can be used to investigate muscle activation patterns during static
+fixations and the control and dynamics of various eye movements. This model can
+be used to calculate metabolic costs of eye movements and also simulate
+different eye disorders, such as different forms of strabismus. Furthermore, it
+can be easily integrated into other human body models to analyze the relation
+between vestibular system and eye movement by simulating their biomechanics in
+three dimensions and investigating the function and neural control of the
+\gls{em}.
+
+Eye movements are produced through the activation of six \gls{em}. Clinical
+trials have provided a profound knowledge of how the \gls{em} rotate the globe,
+the resistive tension to this rotation and the length-tension relationship of
+the muscles. Various computational models of the extraocular muscle and orbital
+mechanics have been proposed, which provide insight and scientific bases for
+oculomotor biomechanics, control of eye movement and binocular
+misalignment. These models focus on the realism of muscle behavior and they were
+based on the viscoelastic properties and physiological data \gls{em}.
+
+The first 3D biomechanical model was developed by \hl{Robinson (1975)}, who
+simplified the formulation by only considering the elasticity of the \gls{em}
+and ignoring the dynamics. The model incorporates anatomically realistic muscle
+paths and empirical innervation-length-tension relationships. To study the
+neural control of rapid saccadic movement, models using anatomical and
+mechanical properties of \gls{em} have been developed. They take account that
+the actual force is a complex nonlinear function of the muscle length, velocity,
+and innervation~\cite{Thelen2003, Millard2013}. Such models have the advantage
+of supporting 3D dynamic simulations and have been used to analyze neural
+controllers and the pulley hypothesis.
+
+Here, we introduce a detailed 3D biomechanical model, based on the work of
+\hl{Iskander} of the human eye which can be used for dynamic simulation. The
+model is based on the biomechanical simulator
+\texttt{OpenSim}~\cite{Delp2007}. This is an open-source software that can
+provide the flexibility of in depth parameterization of the model. The
+architecture and dynamic muscle properties are based on physiological and
+kinematic measurements of the human eye. The model incorporates an eye-globe,
+orbital suspension tissues and six muscles with their connective tissues to test
+the passive pulley hypothesis. Furthermore, with this model we are able to
+assess the excitation and activation patterns for a variety of targets by
+applying a closed-loop fixation controller that drives the model to perform
+saccadic movements in a forward dynamic manner. The controller minimizes the
+error between the desired trajectory (reference input) and the predicted
+movement (calculated output).
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
@@
 muscles, and connective tissues. The size of an emmetropic human adult eye is
-approximately 24.2 mm (transverse, horizontal), 23.7 mm (sagittal, vertical),
-22.0–24.8 mm (axial, anteroposterior) with no significant difference between
+approximately 0.0242 m (transverse, horizontal), 0.0237 mm (sagittal, vertical),
+0.022–0.0248 m (axial, anteroposterior) with no significant difference between
 sexes and age groups. In the transverse diameter, the eyeball size may vary from
-21 mm to 27 mm. Thus, it can be approximated by a solid sphere with 12 mm
-radius. The eyeball was constructed in Blender, an open-source software 3D
-creation Software. We used a spherical mesh with 32 segments and 12 rings, to
+0.021 m to 0.027 m. Thus, it can be approximated by a solid sphere with 0.012 m
+radius. The eyeball was constructed in \texttt{Blender}, an open-source software
+3D creation software. We used a spherical mesh with 32 segments and 12 rings, to
 construct the vitreous humor (body) as solid sphere and a conical plate to
-construct the cornea. The weight of an average human eye is 7.5 g and the
+construct the cornea. The weight of an average human eye is 0.0075 kg and the
 moment of inertia can be calculated similarly as in the case of a spherical
-homogenous and isotropic object with radius 12 mm ($\mat{I} = 2/5 m r^2$ at the
-center of mass).
+homogeneous and isotropic object with radius 0.012 m ($\mat{I} = 2/5 m r^2$ at
+the center of mass).
 
@@
 
+The six \gls{em}, including four rectus muscles and two oblique muscles, are
+controlled by the cranial nerves so as to track a visual target and to stabilize
+the image of the object. The \gls{lr} and \gls{mr} muscles form an antagonistic
+pair to produce horizontal eye movements. The \gls{sr} and \gls{ir} muscles form
+the vertical antagonist pair, which mainly controls vertical eye movement and
+also affects rotation about the horizontal plane and the line of sight
+(secondary action) due to insertion positions and the path of the muscles. The
+\gls{so} muscle passes through the cartilaginous trochlea attached to the
+orbital wall, which reflects the \gls{so} path by 51 deg. The \gls{io} muscle
+originates from the orbital wall anteroinferior to the globe center and inserts
+on the sclera posterior to the globe equator. The primary actions of \gls{so}
+and \gls{io} cause rotation of the globe around the visual axis and vertical
+movement.
+
+We have chosen the passive pulley model for our extraocular model in order to
+keep it simple and provide faster simulation speed. \autoref{tab:muscle-path}
+shows the positions of muscle pulleys, as well as the origin and insertions
+points of the \gls{em} onto the eye, with respect the center of the globe. These
+data were acquired from \hl{Iskander 2018} which were based on physiological
+measurements, but in our case, they were slightly modified to be located outside
+of the globe radius in order to not penetrate the eye globe. Since no position
+was documented for the origin of the \gls{so}, a point close to the origins of
+the rectus muscles was chosen to match the fiber length in the primary position
+of the \gls{so} muscle.
+
 \begin{table}[h]
-  \caption{TODO}\label{tab:muscle-path}
+  \caption{Muscle path for the six \gls{em} (dimensions are given in
+    meters).}\label{tab:muscle-path}
   \begin{tabular}{@{}cccccccccc@{}}
@@
     & \multicolumn{3}{c}{\textbf{Pulley}}
-    & \multicolumn{3}{c}{\textbf{Insertion}}
-    \\ \midrule
+    & \multicolumn{3}{c}{\textbf{Insertion}} \\
+    \midrule
     & \textit{\textbf{Ox}} & \textit{\textbf{Oy}} & \textit{\textbf{Oz}}
     & \textit{\textbf{Px}} & \textit{\textbf{Py}} & \textit{\textbf{Pz}}
-    & \textit{\textbf{Ix}} & \textit{\textbf{Iy}} & \textit{\textbf{Iz}}
-    \\ \midrule
-    LR & -0.034 & 0.0006 & -0.013 & -0.0102 & 0.0003 & 0.012 & 0.0065 & 0 & 0.0101 \\
-    MR & -0.030 & 0.0006 & -0.017 & -0.0053 & 0.00014 & -0.0146 & 0.0088 & 0 & -0.0096 \\
-    SR & -0.0317 & 0.0036 & -0.016 & -0.0092 & 0.012 & -0.002 & 0.0076 & 0.0104 & 0 \\
-    IR & -0.0317 & -0.0024 & -0.016 & -0.0042 & -0.0128 & -0.0042 & 0.00805 & -0.0102 & 0 \\
-    SO & 0.0082 & 0.0122 & -0.0152 & -0.030834 & 0.001145 & -0.01644 & 0.0044 & 0.011 & 0.0029 \\
-    IS & 0.0113 & -0.0154 & -0.0111 & -0.00718 & -0.0135 & 0 & -0.008 & 0 & 0.009 \\
+    & \textit{\textbf{Ix}} & \textit{\textbf{Iy}} & \textit{\textbf{Iz}} \\
+    \midrule
+    \gls{lr} & -0.034 & 0.0006 & -0.013 & -0.0102 & 0.0003 & 0.012 & 0.0065 & 0 & 0.0101 \\
+    \gls{mr} & -0.030 & 0.0006 & -0.017 & -0.0053 & 0.00014 & -0.0146 & 0.0088 & 0 & -0.0096 \\
+    \gls{sr} & -0.0317 & 0.0036 & -0.016 & -0.0092 & 0.012 & -0.002 & 0.0076 & 0.0104 & 0 \\
+    \gls{ir} & -0.0317 & -0.0024 & -0.016 & -0.0042 & -0.0128 & -0.0042 & 0.00805 & -0.0102 & 0 \\
+    \gls{so} & 0.0082 & 0.0122 & -0.0152 & -0.030834 & 0.001145 & -0.01644 & 0.0044 & 0.011 & 0.0029 \\
+    \gls{io} & 0.0113 & -0.0154 & -0.0111 & -0.00718 & -0.0135 & 0 & -0.008 & 0 & 0.009 \\
     \bottomrule
@@
 \end{table}
+
+The Millard muscle model~\cite{Millard2013} has been adopted for the modeling of
+the \gls{em}, which allows to manually fit force-generation dynamics. The six
+\gls{em} were modeled using the rigid tendon assumption that ignores the
+elasticity of the tendon. This means that the series element of the muscle model
+is not included (the tendon length $l^T$ is equal to the tendon slack length
+$l_s^T$). This assumption is valid when the ratio of the tendon length to the
+muscle length is less or equal to one, as the in the case of all
+\gls{em}. \gls{em} are considered parallel-fibered muscles, so the pennation
+angle is zero ($\alpha = 0$). Maximum isometric force $f_o^M$, optimal fiber
+length $l_o^M$ and tendon length $l^T$ are presented
+in~\autoref{tab:muscle-path} and are based on \hl{Iskander 2018}.
+
+The active and passive \gls{fl} curves for the \gls{em} differ from the ones
+that of the skeletal muscles. As shown in \autoref{fig:millard-curves} can be
+adapted so as to fit the experimental data available for the gls{lr} muscle. The
+following values where used for the active \gls{fl} curve:
+
+\begin{itemize}
+\item min norm active fiber length: 0.55
+\item transition norm fiver length: 0.7
+\item max norm active fiver length: 1.8
+\item shallow ascending slope: 2.4
+\item minimum value: 0.0
+\end{itemize}
+%
+and for the passive \gls{fl} curve accordingly:
+
+\begin{itemize}
+\item strain at zero force: -0.18
+\item strain at one norm force: 0.4
+\end{itemize}
+
+\begin{figure}[!ht]
+  \subfloat[Active \gls{fl} curve]{\includegraphics[width=0.5\textwidth,
+    keepaspectratio]{active-force-length-curve.png}\label{fig:active-force-length-curve}}
+  \subfloat[Passive \gls{fl} curve]{\includegraphics[width=0.5\textwidth,
+    keepaspectratio]{passive-force-length-curve.png}\label{fig:passive-force-length-curve}}
+  \caption{The active and passive \gls{fl} curve definition of the Millard
+    muscle model as implemented in \texttt{OpenSim}.}\label{fig:millard-curves}
+\end{figure}
+
+The parameters that describe the above relationships were chosen to fit the
+curves found in \hl{Iskander 2018}. In the plots below, we try to match the
+\gls{fl} relationships at maximum activation of the lateral rectus muscle. This
+represents the first part of testing the fidelity of the model. Due to lack of
+data describing the other muscles’ \gls{fl} relationships, we used the
+parameters found describing the normalized curves of active and passive \gls{fl}
+relationships of the gls{lr} for the other \gls{em} as well.
+
+\gls{em} have a higher fraction of fast twitch fibers and thus different
+\gls{fv} behavior, due to different structures compared to skeletal muscles. The
+default Millard \gls{fv} curve was used for the six \gls{em}, since the behavior
+of the selected muscle model depends mainly on the maximum contraction velocity
+$v^{\text{max}}$. The maximum muscle contraction velocity is tuned so as to
+match the peak velocity of saccadic eye movement 15.7 rad / s. Following this
+definition, the maximum muscle contraction velocity is given in optimal fiber
+length per seconds and it is thus different for each \gls{em} since their
+optimal fiber length is different. Also, because of different structure of
+neural control of eye movements, activation and deactivation delays are lower
+than in skeletal muscles, at about 5 ms.
+
+\begin{table}[h]
+  \caption{Millard muscle parameters for the
+    \gls{em}.}\label{tab:muscle-parameters}
+  \begin{tabular}{@{}cccccccccc@{}}
+    \toprule
+    \thead{Muscle}
+    & \thead{Maximum Isometric \\ Force (N)}
+    & \thead{Optimal Fiber \\ Length (m)}
+    & \thead{Tendon Slack \\ Length (m)}
+    & \thead{Maximum Contraction \\ Velocity (m / s)} \\
+    \midrule
+    \gls{lr} & 1.4710 & 0.04898 & 0.0084 & 3.8483 \\
+    \gls{mr} & 1.5740 & 0.04084 & 0.0038 & 4.6155 \\
+    \gls{sr} & 1.1768 & 0.04487 & 0.0054 & 4.2009 \\
+    \gls{ir} & 1.4269 & 0.04549 & 0.0048 & 4.1437 \\
+    \gls{so} & 0.6031 & 0.03956 & 0.0265 & 4.7648 \\
+    \gls{io} & 0.5590 & 0.04110 & 0.0015 & 3.5863 \\
+    \bottomrule
+  \end{tabular}
+\end{table}
+
+Wrapping objects help to simulate the proper dynamics. Using wrapping objects
+implemented in the \texttt{OpenSim} allows us to model realistic force
+directions as the muscle runs over the eyeball. From the point of origin to the
+pulley point, the force exerted by the contraction of the muscles is aligned
+along one straight line. But, from the pulley point to the insertion point, the
+force is distributed along the surface of a sphere. Two separate wrapping
+spheres for the rectus muscles and the oblique muscles were created, to avoid
+abnormal changes on the \gls{fl} curve as the eyeball rotates in the three
+directions.
 
@@
 % 
-where, $\vec{f}_t$ represents the passive tissue forces, $k_p= 0.33$ g / deg,
-$k_p= 1.56$ g / (deg^3) (\hl{TODO:convert to $Nm / rad$}) and $k_v= 10^{-4}$ Nm
-s / rad the constants and $\vec{\dot{q}} \inr{3}$ the rotational coordinates of
-the model. These forces serve the eye’s stabilization, and are modeled using
-\texttt{OpenSim}'s expression based coordinate force.
-% 0.33 * 9.8066500286389 * 10**-3 / (numpy.pi / 180) = 0.18542028707494282
-% 1.56 * 9.8066500286389 * 10**-3 / (numpy.pi / 180) ** 3 = 2877.4856893811248
+where, $\vec{f}_t$ represents the passive tissue forces, $k_p= 0.002225$ N m /
+rad, $k_c= 34.5297$ N m / (rad^3) and $k_v= 0.002$ N m s / rad the constants and
+$\vec{\dot{q}} \inr{3}$ the rotational coordinates of the model. These forces
+serve the eye’s stabilization, and are modeled using \texttt{OpenSim}'s
+expression based coordinate force.
+% 0.33 * 9.8066500286389 * 10**-3 / (numpy.pi / 180) = 0.18542028707494282 * 0.012
+% 1.56 * 9.8066500286389 * 10**-3 / (numpy.pi / 180) ** 3 = 2877.4856893811248 * 0.012
 
@@
 validated. In our study, verification in oculomotor models was performed by
-comparing the force-length characteristic curves of the modeled \gls{em} to
+comparing the \gls{fl} characteristic curves of the modeled \gls{em} to
 published data. These data include only the lateral rectus characteristic
 curves. However, we can make a safe assumption that the other muscles have
-similar properties. More characteristic curves showing the force-length
-relationship of the other \gls{em}, as well as the changes in muscle length with
-rotation in all directions, are presented in the end of this report. Further
-verification can be done by comparing the model’s joint forces produced during
-horizontal movement with clinically collected data, but these data are not
-available.  Validation was performed by simulating and analyzing synthesized eye
-movements while ensuring that Listing Law was obeyed by maintaining zero torsion
-in secondary gaze positions, and by testing if the assessed innervations are
+similar properties. More characteristic curves showing the \gls{fl} relationship
+of the other \gls{em}, as well as the changes in muscle length with rotation in
+all directions, are presented in the end of this report. Further verification
+can be done by comparing the model’s joint forces produced during horizontal
+movement with clinically collected data, but these data are not available.
+Validation was performed by simulating and analyzing synthesized eye movements
+while ensuring that Listing Law was obeyed by maintaining zero torsion in
+secondary gaze positions, and by testing if the assessed innervations are
 sufficient in fixating the eye at a desired position.
@@
 
-We manage to construct a complete ocular model that represents the ocular
-motility of a normal human eye. The model can be used to drive simulations of
-different eye movement systems and be derive some inference about the excitation
-and activation patterns. The verification and validation of the model showed
-that it matches the data found in literature and it is able to produce
-synthetized movements. We even showed that the simulation results produced by
-static optimization and forward dynamics in OpenSim are less satisfactory than
-the proposed solution where we used a PD controller to minimize the tracking
-error between the desired and estimated trajectory of saccadic movements. The
-desired kinematics were tracked within a maximum \gls{rmse} of and in horizontal
-and vertical saccades respectively. With the application of the controller, a
-very rapid instantaneous acceleration and a constant velocity to sustain clear
-vision is attained. The produced activation levels were in accordance with the
+A realistic ocular model that represents the ocular motility of a normal human
+eye was presented. The model can be used to simulate different saccadic
+movements and obtain the muscle activations required to track the movement. The
+model was validated against experimental measured data and is able to synthesize
+saccadic movements reasonably well.
+
+We even showed that the simulation results produced by static optimization and
+forward dynamics in OpenSim are less satisfactory than the proposed solution
+where we used a \gls{pd} controller to minimize the tracking error between the
+desired and estimated trajectory of saccadic movements. The desired kinematics
+were tracked within a maximum \gls{rmse} of and in horizontal and vertical
+saccades respectively. With the application of the controller, a very rapid
+instantaneous acceleration and a constant velocity to sustain clear vision is
+attained. The produced activation levels were in accordance with the
 descriptions found in the bibliography and the highest activation levels were
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% \bibliography{mylibrary}
-% \addcontentsline{toc}{chapter}{Bibliography}
+
+\bibliography{mylibrary}
+\addcontentsline{toc}{chapter}{Bibliography}
 

### 2018-07-09 doc/manuscript.tex
fix fixation controller

@@
 \usepackage[pdfpagelabels,pdfusetitle,colorlinks=true,pdfborder={0 0
-  0}]{hyperref}
+  0},pdfauthor={Dimitar Stanev}]{hyperref}
 \usepackage{natbib}
@@
 % \newacronym{so}{SO}{Static Optimization}
+\newacronym{fd}{FD}{Forward Dynamics}
 \newacronym{pd}{PD}{Proportional Derivative}
@@
 \renewcommand*{\vec}[1]{\bm{#1}}
-\newcommand{\pd}[2]{\frac{\partial#1}{\partial#2}}
-\newcommand{\subspace}[1]{\varmathbb{#1}}
-\newcommand{\rs}[1]{\subspace{R} (#1)}
-\newcommand{\cs}[1]{\subspace{C} (#1)}
-\newcommand{\ns}[1]{\subspace{N} (#1)}
-\newcommand{\fs}[1]{\subspace{F}}
-\newcommand{\ms}[1]{\subspace{M}}
 \newcommand{\R}[1]{\mathfrak{R}^{#1}}
 \newcommand{\inr}[1]{\in\R{#1}}
-\newcommand{\TP}[1]{\mat{T}_{#1}}     % tall projection
-\newcommand{\FP}[1]{\mat{F}_{#1}}     % fat projection
-\newcommand{\NT}[1]{\mat{N}_{\TP{#1}}}
-\newcommand{\N}[1]{\mat{N}_{#1}}
-\newcommand{\dNT}[1]{\mat{\dot{N}}_{\TP{#1}}}
-\newcommand{\NF}[1]{\mat{N}_{\FP{#1}}}
-\newcommand{\dNF}[1]{\mat{\dot{N}}_{\FP{#1}}}
-\newcommand{\mc}[1]{\mathcal{#1}}
 
-% mathematical proofs
-\newtheorem{theorem}{Theorem}[chapter]
-\newtheorem{corollary}{Corollary}[chapter]
-\newcommand{\corollaryautorefname}{Corollary}
-\newtheorem{proposition}{Proposition}[chapter]
-\newcommand{\propositionautorefname}{Proposition}
-\newtheorem{remark}{Remark}[chapter]
-\newcommand{\remarkautorefname}{Remark}
-\newtheorem{definition}{Definition}[chapter]
-\newcommand{\definitionautorefname}{Definition}
 \newcommand{\subfigureautorefname}{Figure} % subfloat
@@
 \def\equationautorefname~#1\null{Equation #1\null}
-
-% \renewcommand{\baselinestretch}{1.5} % line stretch
 
@@
 gaze. Studying human eye movement has significant implications for improving our
-understanding of the oculomotor system and treating visuomotor disorders. Over
-the years, biomechanic simulation has provided an analysis tool of different
-human movements, especially gait. This analysis can be extended to visual tasks
-by analyzing the mechanisms of ocular motility and providing a realistic ocular
-model that can be used to investigate muscle activation patterns during static
-fixations and the control and dynamics of various eye movements. This model can
-be used to calculate metabolic costs of eye movements and also simulate
-different eye disorders, such as different forms of strabismus. Furthermore, it
-can be easily integrated into other human body models to analyze the relation
-between vestibular system and eye movement by simulating their biomechanics in
-three dimensions and investigating the function and neural control of the
-\gls{em}.
+understanding of the oculomotor system and treating vasomotor visuomotor disorders. Over
+the past decades, biomechanics simulation has provided the means to analyze
+different human movements. The same principles can be used to analyze visual
+tasks by modeling the musculoskeletal properties of the oculomotor
+system. Consequently, this model can be used to investigate muscle activation
+patterns during static fixation, analyze the dynamics of various eye movements,
+calculate metabolic costs and simulate different eye disorders, such as
+different forms of strabismus. Last but not least, it can be easily integrated
+with available full body models in order to analyze the relation between the
+vestibular and oculomotor systems.
 
-Eye movements are produced through the activation of six \gls{em}. Clinical
-trials have provided a profound knowledge of how the \gls{em} rotate the globe,
-the resistive tension to this rotation and the length-tension relationship of
-the muscles. Various computational models of the extraocular muscle and orbital
-mechanics have been proposed, which provide insight and scientific bases for
-oculomotor biomechanics, control of eye movement and binocular
-misalignment. These models focus on the realism of muscle behavior and they were
-based on the viscoelastic properties and physiological data \gls{em}.
+Eye movements are a result of the coordination of the six \gls{em}. Clinical
+trials have provided a profound knowledge of how the \gls{em} act on the eye
+globe, the resistive tension of the surrounding tissues and the length-tension
+relationship of the muscles. Various computational models of the extraocular
+muscles and orbital mechanics have been proposed, which provide insight and
+scientific bases for oculomotor biomechanics, control of eye movement and
+binocular misalignment. These models focus on the realism of muscle behavior and
+they were based on the viscoelastic properties and physiological data \gls{em}.
 
-The first 3D biomechanical model was developed by \hl{Robinson (1975)}, who
-simplified the formulation by only considering the elasticity of the \gls{em}
-and ignoring the dynamics. The model incorporates anatomically realistic muscle
-paths and empirical innervation-length-tension relationships. To study the
-neural control of rapid saccadic movement, models using anatomical and
-mechanical properties of \gls{em} have been developed. They take account that
-the actual force is a complex nonlinear function of the muscle length, velocity,
-and innervation~\cite{Thelen2003, Millard2013}. Such models have the advantage
-of supporting 3D dynamic simulations and have been used to analyze neural
-controllers and the pulley hypothesis.
+The first 3D biomechanical model was developed by~\cite{Robinson1964a,
+  Robinson1969}, who simplified the formulation by considering the elasticity of
+the \gls{em} ignoring their dynamics. The model incorporates anatomically
+realistic muscle paths and empirical innervation-length-tension
+relationships. To study the neural control of rapid saccadic movement, models
+using anatomical and mechanical properties of \gls{em} have been developed. They
+take account that the actual force is a complex nonlinear function of the muscle
+length, velocity, and innervation~\cite{Thelen2003, Millard2013}. Such models
+have the advantage of supporting 3D dynamic simulations and have been used to
+analyze neural controllers and the pulley hypothesis.
 
-Here, we introduce a detailed 3D biomechanical model, based on the work of
-\hl{Iskander} of the human eye which can be used for dynamic simulation. The
-model is based on the biomechanical simulator
-\texttt{OpenSim}~\cite{Delp2007}. This is an open-source software that can
-provide the flexibility of in depth parameterization of the model. The
-architecture and dynamic muscle properties are based on physiological and
-kinematic measurements of the human eye. The model incorporates an eye-globe,
-orbital suspension tissues and six muscles with their connective tissues to test
-the passive pulley hypothesis. Furthermore, with this model we are able to
-assess the excitation and activation patterns for a variety of targets by
-applying a closed-loop fixation controller that drives the model to perform
-saccadic movements in a forward dynamic manner. The controller minimizes the
-error between the desired trajectory (reference input) and the predicted
-movement (calculated output).
+We introduce a detailed open-source 3D biomechanical model of the human eye,
+which can be used for dynamic simulation\footnote{SimTK project:
+  \url{https://simtk.org/projects/eye}}, based on the data reported
+in~\cite{Iskander2018}. The model is implemented in
+\texttt{OpenSim}~\cite{Delp2007}, which is an open-source framework for modeling
+and simulation of musculoskeletal systems. The architecture and dynamic muscle
+properties are based on physiological and kinematic measurements of the human
+eye. The model incorporates an eye-globe, orbital suspension tissues and six
+\gls{em}. Using this model we are able to assess the excitation and activation
+patterns for a variety of targets by applying a closed-loop fixation controller
+that drives the model to perform saccadic movements in a \gls{fd} manner. The
+controller minimizes the error between the desired saccadic trajectory
+(reference input) and the predicted movement (calculated output).
 
@@
 
-We have chosen the passive pulley model for our extraocular model in order to
-keep it simple and provide faster simulation speed. \autoref{tab:muscle-path}
-shows the positions of muscle pulleys, as well as the origin and insertions
-points of the \gls{em} onto the eye, with respect the center of the globe. These
-data were acquired from \hl{Iskander 2018} which were based on physiological
-measurements, but in our case, they were slightly modified to be located outside
-of the globe radius in order to not penetrate the eye globe. Since no position
-was documented for the origin of the \gls{so}, a point close to the origins of
-the rectus muscles was chosen to match the fiber length in the primary position
-of the \gls{so} muscle.
+The model relies on the passive pulley assumption in order to keep it simple and
+provide faster simulation speed. \autoref{tab:muscle-path} shows the positions
+of muscle pulleys, as well as the origin and insertion points of the \gls{em},
+defined in the local body coordinates of the globe. The data is based on
+physiological measurements~\cite{Iskander2018}, with some minor adaptation so as
+to prevent penetration into the eye globe. Since no position was documented for
+the origin of the \gls{so}, a point close to the origins of the rectus muscles
+was chosen to match the fiber length in the primary position of the \gls{so}
+muscle.
 
@@
 The Millard muscle model~\cite{Millard2013} has been adopted for the modeling of
-the \gls{em}, which allows to manually fit force-generation dynamics. The six
-\gls{em} were modeled using the rigid tendon assumption that ignores the
-elasticity of the tendon. This means that the series element of the muscle model
-is not included (the tendon length $l^T$ is equal to the tendon slack length
-$l_s^T$). This assumption is valid when the ratio of the tendon length to the
-muscle length is less or equal to one, as the in the case of all
+the \gls{em}, which allows to manually fit the reported data for the \gls{fl}
+curves. The muscles were modeled using the rigid tendon assumption that ignores
+the elasticity of the tendon. This means that the series element of the muscle
+model is not included (the tendon length $l^T$ is equal to the tendon slack
+length $l_s^T$). This assumption is valid when the ratio of the tendon length to
+the muscle length is less or equal to one, as the in the case of all
 \gls{em}. \gls{em} are considered parallel-fibered muscles, so the pennation
@@
 length $l_o^M$ and tendon length $l^T$ are presented
-in~\autoref{tab:muscle-path} and are based on \hl{Iskander 2018}.
+in~\autoref{tab:muscle-path}.
 
-The active and passive \gls{fl} curves for the \gls{em} differ from the ones
-that of the skeletal muscles. As shown in \autoref{fig:millard-curves} can be
-adapted so as to fit the experimental data available for the gls{lr} muscle. The
-following values where used for the active \gls{fl} curve:
+The active and passive \gls{fl} curves for the \gls{em} differ from that of a
+skeletal muscle. As shown in \autoref{fig:millard-curves}, we can fine-tune
+these curves so as to fit the experimental data available for the gls{lr}
+muscle. The following values where used for the active \gls{fl} curve:
 
@@
 The parameters that describe the above relationships were chosen to fit the
-curves found in \hl{Iskander 2018}. In the plots below, we try to match the
-\gls{fl} relationships at maximum activation of the lateral rectus muscle. This
-represents the first part of testing the fidelity of the model. Due to lack of
-data describing the other muscles’ \gls{fl} relationships, we used the
-parameters found describing the normalized curves of active and passive \gls{fl}
-relationships of the gls{lr} for the other \gls{em} as well.
+curves reported in~\cite{Iskander2018}, by matching the \gls{fl} relationship at
+maximum activation of the \gls{lr} muscle. This represents the first part of
+testing the fidelity of the model. Due to lack of data describing the other
+muscles’ \gls{fl} relationships, we used the parameters found describing the
+normalized curves of active and passive \gls{fl} relationships of the gls{lr}
+for the other \gls{em} as well.
 
 \gls{em} have a higher fraction of fast twitch fibers and thus different
-\gls{fv} behavior, due to different structures compared to skeletal muscles. The
-default Millard \gls{fv} curve was used for the six \gls{em}, since the behavior
-of the selected muscle model depends mainly on the maximum contraction velocity
-$v^{\text{max}}$. The maximum muscle contraction velocity is tuned so as to
-match the peak velocity of saccadic eye movement 15.7 rad / s. Following this
-definition, the maximum muscle contraction velocity is given in optimal fiber
-length per seconds and it is thus different for each \gls{em} since their
-optimal fiber length is different. Also, because of different structure of
-neural control of eye movements, activation and deactivation delays are lower
-than in skeletal muscles, at about 5 ms.
+\gls{fv} behavior, due to different structures compared to skeletal
+muscles. Despite that, the default Millard \gls{fv} curve was used for the six
+\gls{em}, since the behavior of the selected muscle model depends mainly on the
+maximum contraction velocity $v^{\text{max}}$. The maximum muscle contraction
+velocity is tuned so as to match the peak velocity of saccadic eye movement 15.7
+rad / s. Following this definition, the maximum muscle contraction velocity is
+given in optimal fiber length per seconds and it is thus different for each
+\gls{em}, as their optimal fiber length is different. Furthermore, because of
+the different structure of neural control of the eyes, activation and
+deactivation delays ($\tau_d = 5$ ms) are lower than in skeletal
+muscles. Finally, two separate wrapping spheres for the rectus muscles and the
+oblique muscles were created, to avoid abnormal changes on the \gls{fl} curve as
+the eyeball rotates in the three directions.
 
@@
 \end{table}
-
-Wrapping objects help to simulate the proper dynamics. Using wrapping objects
-implemented in the \texttt{OpenSim} allows us to model realistic force
-directions as the muscle runs over the eyeball. From the point of origin to the
-pulley point, the force exerted by the contraction of the muscles is aligned
-along one straight line. But, from the pulley point to the insertion point, the
-force is distributed along the surface of a sphere. Two separate wrapping
-spheres for the rectus muscles and the oblique muscles were created, to avoid
-abnormal changes on the \gls{fl} curve as the eyeball rotates in the three
-directions.
 

### 2018-07-09 source_code/src/fixation_controller_plugin/FixationController.cpp

@@ -22,8 +22,9 @@ using namespace SimTK;
  */
 Vec3 sigmoid(double t, double t0, double A, double B) {
     return Vec3(A * (tanh(B * (t - t0)) + 1) / 2,
-                A * B * (1 - pow(tanh(B * (t - t0)), 2)) / 2,
-                A * pow(B, 2) * tanh(B * (t - t0)) * (1 - pow(tanh(B * (t - t0)), 2)));
+                -(A * B * (pow(tanh(B * (t - t0)), 2) - 1)) / 2,
+                A * pow(B, 2) * tanh(B * (t - t0)) * (pow(tanh(B * (t - t0)), 2) - 1)
+    );
 }
 
 FixationController::FixationController() : Controller() {
@@ -39,8 +40,8 @@ void FixationController::constructProperties() {
     constructProperty_kdV(1.5);
     constructProperty_kpT(100);
     constructProperty_kdT(0.5);
-    constructProperty_saccade_onset(0);
-    constructProperty_saccade_velocity(200);
+    constructProperty_saccade_onset(0.5);
+    constructProperty_saccade_velocity(100);
 }
 
 void FixationController::computeControls(const State& s, Vector& controls) const {
@@ -60,20 +61,23 @@ void FixationController::computeControls(const State& s, Vector& controls) const
     const auto& inf_oblq = _model->getMuscles().get("r_Inferior_Oblique");
 
     // Compute the desired position, velocity and acceleration
+    auto thetaH = convertDegreesToRadians(get_thetaH());
+    auto thetaV = convertDegreesToRadians(get_thetaV());
+    auto desSaccadeVelocity = convertDegreesToRadians(get_saccade_velocity());
     double xdes = 0;
     double xdesv = 0;
     double xdesa = 0;
     auto sigmY = sigmoid(t,
                          get_saccade_onset(),
-                         convertDegreesToRadians(get_thetaH()),
-                         get_saccade_velocity());
+                         thetaH,
+                         2 * desSaccadeVelocity / abs(thetaH));
     double ydes = 0 + sigmY[0];
     double ydesv = sigmY[1];
     double ydesa = sigmY[2];
     auto sigmZ = sigmoid(t,
                          get_saccade_onset(),
-                         convertDegreesToRadians(get_thetaV()),
-                         get_saccade_velocity());
+                         thetaV,
+                         2 * desSaccadeVelocity / abs(thetaV));
[…]

### 2018-07-09 source_code/src/fixation_controller_plugin/FixationController.h

@@ -33,7 +33,7 @@ namespace OpenSim {
         OpenSim_DECLARE_PROPERTY(kpT, double, "torsion position tracking gain");
         OpenSim_DECLARE_PROPERTY(kdT, double, "torsion velocity tracking gain");
         OpenSim_DECLARE_PROPERTY(saccade_onset, double, "saccade onset (s)");
-        OpenSim_DECLARE_PROPERTY(saccade_velocity, double, "saccade velocity (in rad/s^2)");
+        OpenSim_DECLARE_PROPERTY(saccade_velocity, double, "desired saccade velocity (deg / s)");
         /** Constructor. */
         FixationController();
         /** Distractor. */

### 2018-07-09 source_code/src/include/Settings.h

@@ -3,6 +3,6 @@
 
 #include <string>
 
-std::string MODEL_DIR = "D:/Desktop/upat_eye_model/source_code/../model/";
+std::string MODEL_DIR = "C:/Users/mitkof6/Desktop/upat_eye_model/source_code/../model/";
 
 #endif

### 2018-07-09 source_code/src/simulation/Simulation.cpp

@@ -63,7 +63,6 @@ void addExpressionCoordinateForce(Model* model) {
 void simulateModel() {
     Model model(MODEL_DIR + "UPAT_Eye_Model_Passive_Pulleys_v2.osim");
     model.setName("UPAT_Eye_Model_Passive_Pulleys_v3");
-    // model.setUseVisualizer(true);
 
     // Add expression based coordinate force
     addExpressionCoordinateForce(&model);
@@ -80,21 +79,14 @@ void simulateModel() {
     controller->set_kpT(100);
     controller->set_kdT(0.5);
     controller->set_saccade_onset(0.5);
-    controller->set_saccade_velocity(200);
+    controller->set_saccade_velocity(100); // deg / s
     model.addController(controller);
 
     // Build and initialize model
     auto& state = model.initSystem();
     model.equilibrateMuscles(state);
-    /*model.updCoordinateSet().get("r_eye_add_abd").setValue(state, 0.1);
-      model.updCoordinateSet().get("r_eye_sup_inf").setValue(state, 0.1);*/
 
-      // Visualize model
-      /*model.updMatterSubsystem().setShowDefaultGeometry(true);
-        model.updVisualizer().updSimbodyVisualizer().setBackgroundColor(Vec3(1));
-        model.updVisualizer().updSimbodyVisualizer().drawFrameNow(state);*/
-
-        // Create the integrator and manager for the simulation
+    // Create the integrator and manager for the simulation
     RungeKuttaMersonIntegrator integrator(model.getMultibodySystem());
     Manager manager(model, integrator);
 

### 2018-07-09 model/plot_results.py
visualize results

@@ -1,0 +1,135 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*-
+import numpy as np
+import matplotlib.pyplot as plt
+
+
+def readMotionFile(filename):
+    """Reads OpenSim .sto files.
+
+    Parameters
+    ----------
+    filename: str
+        absolute path to the .sto file
+
+    Returns
+    -------
+    header: list of str
+        the header of the .sto
+    labels: list of str
+        the labels of the columns
+    data: list of lists
+        an array of the data
+
+    """
+
+    if not os.path.exists(filename):
+        print('file do not exists')
+
+    file_id = open(filename, 'r')
+
+    # read header
+    next_line = file_id.readline()
+    header = [next_line]
+    nc = 0
+    nr = 0
+    while not 'endheader' in next_line:
+        if 'datacolumns' in next_line:
+            nc = int(next_line[next_line.index(' ') + 1:len(next_line)])
+        elif 'datarows' in next_line:
+            nr = int(next_line[next_line.index(' ') + 1:len(next_line)])
+        elif 'nColumns' in next_line:
+            nc = int(next_line[next_line.index('=') + 1:len(next_line)])
+        elif 'nRows' in next_line:
+            nr = int(next_line[next_line.index('=') + 1:len(next_line)])
+
+        next_line = file_id.readline()
+        header.append(next_line)
+
+    # process column labels
+    next_line = file_id.readline()
+    if next_line.isspace() == True:
+        next_line = file_id.readline()
+
+    labels = next_line.split()
+
+    # get data
+    data = []
+    for i in range(1, nr + 1):
+        d = [float(x) for x in file_id.readline().split()]
+        data.append(d)
+
+    file_id.close()
+
+    return header, labels, data
+
+
+def index_containing_substring(list_str, pattern):
+    """For a given list of strings finds the index of the element that contains the
+    substring.
+
+    Parameters
+    ----------
+    list_str: list of str
+
+    pattern: str
+         pattern
+
+
+    Returns
+    -------
+    indices: list of int
+         the indices where the pattern matches
+
+    """
+    indices = []
+    for i, s in enumerate(list_str):
+        if pattern in s:
+            indices.append(i)
+
+    return indices
+
[…]

### 2018-07-09 doc/manuscript.tex
prepare a first version of the manuscript

@@
 \newacronym{eoms}{EoMs}{Equations of Motion}
-% \newacronym{so}{SO}{Static Optimization}
 \newacronym{fd}{FD}{Forward Dynamics}
@@
 \newacronym{io}{IO}{Inferior Oblique}
-\newacronym{rmse}{RMSE}{Root Mean Square Error}
 \newacronym{em}{EOMs}{Extraocular Muscles}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
-\title{An Open-Source OpenSim Oculomotor Model for Kinematic and Dynamic
+\title{An Open-Source OpenSim Oculomotor Model for Kinematics and Dynamics
   Simulation}
@@
 \author{Constantinos Filip, Dimitar Stanev\footnote{Electrical and Computer
-    Engineering Department, University of Patras, Greece, Corresponding author:
-    stanev@ece.upatras.gr} and Konstantinos Moustakas}
+    Engineering Department, University of Patras, Greece; Corresponding
+    author: \url{stanev@ece.upatras.gr}}, and Konstantinos Moustakas}
 
@@
 \begin{abstract}
-  bla
+  Studying human eye movement has significant implications for improving our
+  understanding of the oculomotor system and treating various visuomotor
+  disorders. An open-source musculoskeletal model of the human eye, that can be
+  used for kinematics and dynamics analysis, is implemented based on the data
+  reported in literature and made publicly available\footnote{SimTK project:
+    \url{https://simtk.org/projects/eye}}. The model is implemented in
+  \texttt{OpenSim}~\cite{Delp2007}, which is an open-source framework for
+  modeling and simulation of musculoskeletal systems. The calibration of the
+  model parameters is based on physiological measurements of the human
+  eye~\cite{Iskander2018}. The model incorporates an eye globe, orbital
+  suspension tissues and six \gls{em}. The excitation and activation patterns
+  for a variety of targets can be calculated using a closed-loop fixation
+  controller that drives the model to perform saccadic movements in a \gls{fd}
+  manner. The controller minimizes the error between the desired saccadic
+  trajectory and the predicted movement. Consequently, this model enables the
+  investigation muscle activation patterns during static fixation and analyze
+  the dynamics of various eye movements.
 \end{abstract}
@@
 gaze. Studying human eye movement has significant implications for improving our
-understanding of the oculomotor system and treating vasomotor visuomotor disorders. Over
+understanding of the oculomotor system and treating visuomotor disorders. Over
 the past decades, biomechanics simulation has provided the means to analyze
@@
 calculate metabolic costs and simulate different eye disorders, such as
-different forms of strabismus. Last but not least, it can be easily integrated
-with available full body models in order to analyze the relation between the
+different forms of strabismus. Furthermore, it can be easily integrated with
+available full body models in order to analyze the relation between the
 vestibular and oculomotor systems.
 
-Eye movements are a result of the coordination of the six \gls{em}. Clinical
+Eye movements are a generated from a coordination of the six \gls{em}. Clinical
 trials have provided a profound knowledge of how the \gls{em} act on the eye
@@
 realistic muscle paths and empirical innervation-length-tension
-relationships. To study the neural control of rapid saccadic movement, models
-using anatomical and mechanical properties of \gls{em} have been developed. They
-take account that the actual force is a complex nonlinear function of the muscle
-length, velocity, and innervation~\cite{Thelen2003, Millard2013}. Such models
-have the advantage of supporting 3D dynamic simulations and have been used to
-analyze neural controllers and the pulley hypothesis.
-
-We introduce a detailed open-source 3D biomechanical model of the human eye,
-which can be used for dynamic simulation\footnote{SimTK project:
-  \url{https://simtk.org/projects/eye}}, based on the data reported
-in~\cite{Iskander2018}. The model is implemented in
-\texttt{OpenSim}~\cite{Delp2007}, which is an open-source framework for modeling
-and simulation of musculoskeletal systems. The architecture and dynamic muscle
-properties are based on physiological and kinematic measurements of the human
-eye. The model incorporates an eye-globe, orbital suspension tissues and six
-\gls{em}. Using this model we are able to assess the excitation and activation
-patterns for a variety of targets by applying a closed-loop fixation controller
-that drives the model to perform saccadic movements in a \gls{fd} manner. The
-controller minimizes the error between the desired saccadic trajectory
-(reference input) and the predicted movement (calculated output).
+relationships. To study the neural control of rapid saccadic movements, models
+using anatomical and mechanical properties of \gls{em} have been developed by
+accounting for the nonlinear muscle dynamics~\cite{Thelen2003a,
+  Millard2013}. Such models, having the advantage of supporting dynamics
+simulation, are used in conjunction with brain level controllers \hl{seb}.
 
@@
 The orbital plant consists of the globe (eyeball), three pairs of extraocular
-muscles, and connective tissues. The size of an emmetropic human adult eye is
-approximately 0.0242 m (transverse, horizontal), 0.0237 mm (sagittal, vertical),
-0.022–0.0248 m (axial, anteroposterior) with no significant difference between
-sexes and age groups. In the transverse diameter, the eyeball size may vary from
-0.021 m to 0.027 m. Thus, it can be approximated by a solid sphere with 0.012 m
-radius. The eyeball was constructed in \texttt{Blender}, an open-source software
-3D creation software. We used a spherical mesh with 32 segments and 12 rings, to
-construct the vitreous humor (body) as solid sphere and a conical plate to
-construct the cornea. The weight of an average human eye is 0.0075 kg and the
-moment of inertia can be calculated similarly as in the case of a spherical
-homogeneous and isotropic object with radius 0.012 m ($\mat{I} = 2/5 m r^2$ at
-the center of mass).
+muscles and the connective passive tissues. The size of an emmetropic human
+adult eye is approximately 0.0242 m (transverse, horizontal), 0.0237 m
+(sagittal, vertical), 0.022–0.0248 m (axial, anteroposterior) with no
+significant difference between sexes and age groups. In the transverse diameter,
+the eyeball size may vary from 0.021 m to 0.027 m. Thus, it can be approximated
+by a solid sphere with 0.012 m radius. The eyeball was constructed in
+\texttt{Blender}, an open-source software 3D creation software. We used a
+spherical mesh with 32 segments and 12 rings, to construct the vitreous humor
+(body) as solid sphere and a conical plate to construct the cornea. The weight
+of an average human eye is 0.0075 kg and the moment of inertia can be calculated
+similarly as in the case of a spherical homogeneous and isotropic object with
+radius 0.012 m ($\mat{I} = 2/5 m r^2$ at the center of mass).
 
@@
 
-\begin{table}[h]
+\begin{table}[ht]
   \caption{Muscle path for the six \gls{em} (dimensions are given in
@@
 
-\begin{figure}[!ht]
+\begin{figure}[ht]
   \subfloat[Active \gls{fl} curve]{\includegraphics[width=0.5\textwidth,
@@
 
-\begin{table}[h]
+\begin{table}[ht]
   \caption{Millard muscle parameters for the
@@
 where, $\vec{f}_t$ represents the passive tissue forces, $k_p= 0.002225$ N m /
-rad, $k_c= 34.5297$ N m / (rad^3) and $k_v= 0.002$ N m s / rad the constants and
-$\vec{\dot{q}} \inr{3}$ the rotational coordinates of the model. These forces
-serve the eye’s stabilization, and are modeled using \texttt{OpenSim}'s
-expression based coordinate force.
+rad, $k_c= 34.5297$ N m / (rad\textsuperscript{3}) and $k_v= 0.002$ N m s / rad
+the constants and $\vec{\dot{q}} \inr{3}$ the rotational coordinates of the
+model. These forces serve the eye’s stabilization, and are modeled using
+\texttt{OpenSim}'s expression based coordinate force.
+
 % 0.33 * 9.8066500286389 * 10**-3 / (numpy.pi / 180) = 0.18542028707494282 * 0.012
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
-\subsection*{Model Validation}\label{sec:model-validation}
+\subsection*{Fixation Controller}\label{sec:fixation-controller}
 
-To meet good fidelity criteria, a model requires to be verified and
-validated. In our study, verification in oculomotor models was performed by
-comparing the \gls{fl} characteristic curves of the modeled \gls{em} to
-published data. These data include only the lateral rectus characteristic
-curves. However, we can make a safe assumption that the other muscles have
-similar properties. More characteristic curves showing the \gls{fl} relationship
-of the other \gls{em}, as well as the changes in muscle length with rotation in
-all directions, are presented in the end of this report. Further verification
-can be done by comparing the model’s joint forces produced during horizontal
-movement with clinically collected data, but these data are not available.
-Validation was performed by simulating and analyzing synthesized eye movements
-while ensuring that Listing Law was obeyed by maintaining zero torsion in
-secondary gaze positions, and by testing if the assessed innervations are
-sufficient in fixating the eye at a desired position.
+A fixation controller was implemented as a custom \texttt{OpenSim} plugin. The
+parameters of the controller are: the desired horizontal $\theta_H$ and vertical
+$\theta_V$ fixation angles (in degrees), the saccade onset and velocity and the
+gains of \gls{pd} tracking controller ($k_p$ and $k_d $). A sigmoid function is
+used for generating smooth saccade trajectories
+
+\begin{equation}\label{equ:sigmoid}
+  \begin{aligned}
+    \theta_d(t) &= \frac{a}{2} \Big(\tanh(b (t - t_0)) + 1\Big) \\
+    \dot{\theta}_d(t) &= \frac{a b}{2} \Big(\tanh^2(b (t - t_0)) - 1\Big) 
+  \end{aligned}
+\end{equation}
+%
+where $\theta_d(t)$ and $\dot{\theta}_d(t)$ represent the desired orientation
+and velocity at time $t$, $a$ the magnitude of the trajectory, $b$ the slope and
+$t_0$ a time shift constant. Provided a fixation goal $\theta_g$ and a desired
+saccade velocity $\dot{\theta}_g$ the parameters of the sigmoid function are
+defined as follows $a = \theta_g$, $b = 2 \dot{\theta}_g / \theta_g$. The
+\gls{pd} tracking controller has the following form
+
+\begin{equation}\label{equ:pd-controller}
+  \ddot{\theta}(t) = k_p (\theta_d(t) - \theta(t)) + k_d (\dot{\theta}_d(t) -
+  \dot{\theta}(t)) 
+\end{equation}
+
+The sign and magnitude of $\ddot{\theta}(t)$ both for the horizontal and
+vertical coordinates of the fixation target is used to actuate the corresponding
+muscle in order to achieve the desired goal. \autoref{fig:model} presents an
+instance of the model during simulation with the corresponding muscles
+activated. \autoref{fig:simulated-saccade} depicts the simulated coordinates,
+speeds and estimated \gls{em} activations that reproduce the desired saccade
+trajectory.
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=.6\textwidth]{model.png}
+  \caption{Model with a fixation at $\theta_H = -15$ deg $\theta_V = 15$ deg
+    during simulation. Blue denotes low and red high activation
+    values.}\label{fig:model}
+\end{figure}
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=1.\textwidth]{UPAT_Eye_Model_Passive_Pulleys_v3_States[-15][15].pdf}
+  \caption{Simulated saccade response with a fixation at $\theta_H = -15$ deg
+    $\theta_V = 15$ deg. The left subplot represents the simulated generalized
+    coordinates; the middle the coordinate speeds; right the estimated \gls{em}
+    activations.}\label{fig:simulated-saccade}
+\end{figure}
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
-\subsection*{Fixation Controller}\label{sec:fixation-controller}
+\section*{Conclusion}\label{sec:concluison}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
-% \section*{Discussion}\label{sec:discussion}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
-\section*{Conclusion}\label{sec:conclusion}
-
-A realistic ocular model that represents the ocular motility of a normal human
-eye was presented. The model can be used to simulate different saccadic
-movements and obtain the muscle activations required to track the movement. The
-model was validated against experimental measured data and is able to synthesize
-saccadic movements reasonably well.
-
-We even showed that the simulation results produced by static optimization and
-forward dynamics in OpenSim are less satisfactory than the proposed solution
-where we used a \gls{pd} controller to minimize the tracking error between the
-desired and estimated trajectory of saccadic movements. The desired kinematics
-were tracked within a maximum \gls{rmse} of and in horizontal and vertical
-saccades respectively. With the application of the controller, a very rapid
-instantaneous acceleration and a constant velocity to sustain clear vision is
-attained. The produced activation levels were in accordance with the
-descriptions found in the bibliography and the highest activation levels were
-shown only during the time intervals when the main agonist muscle was activated.
+A realistic oculomotor model representing the motility of a normal human eye was
+presented and made publicly available. The parameters of the model were
+calibrated using available experimental measured data. The model can be used for
+kinematics and dynamics analysis or as a tool for obtaining the muscle
+activations that generate a desired saccade, using a closed-loop fixation
+controller in a \gls{fd} manner. There is of course space for further
+improvement, which will enhance the accuracy and the predictability of the
+proposed computational knee model. In this study, we didn't attempt to model the
+muscle pulleys, which vary as a function of the model coordinates. Finally, the
+users should consider validation of the eye model based on the requirements of
+the targeted utility of the model and the variables of interests.
 

### 2018-07-10 doc/manuscript.tex
upd

@@
 parameters of the controller are: the desired horizontal $\theta_H$ and vertical
-$\theta_V$ fixation angles (in degrees), the saccade onset and velocity and the
-gains of \gls{pd} tracking controller ($k_p$ and $k_d $). A sigmoid function is
+$\theta_V$ fixation angles (in degrees), the saccade onset and velocity, and the
+gains of \gls{pd} tracking controller ($k_p$, $k_d $). A sigmoid function is
 used for generating smooth saccade trajectories
@@
   \ddot{\theta}(t) = k_p (\theta_d(t) - \theta(t)) + k_d (\dot{\theta}_d(t) -
-  \dot{\theta}(t)) 
+  \dot{\theta}(t)).
 \end{equation}
@@
 vertical coordinates of the fixation target is used to actuate the corresponding
-muscle in order to achieve the desired goal. \autoref{fig:model} presents an
+muscles in order to achieve the desired goal. \autoref{fig:model} presents an
 instance of the model during simulation with the corresponding muscles
@@
 improvement, which will enhance the accuracy and the predictability of the
-proposed computational knee model. In this study, we didn't attempt to model the
-muscle pulleys, which vary as a function of the model coordinates. Finally, the
-users should consider validation of the eye model based on the requirements of
-the targeted utility of the model and the variables of interests.
+proposed computational eye model. In this study, we didn't attempt to model the
+muscle pulleys, which vary as a function of the model coordinates. Furthermore,
+the users should consider validation of the eye model based on the requirements
+of the targeted utility of the model and the variables of interests.
 
@@
 
-\bibliography{mylibrary}
+\bibliography*{mylibrary}
 \addcontentsline{toc}{chapter}{Bibliography}

### 2018-07-10 doc/manuscript.tex
update text

@@
 Rapid and accurate eye movements are crucial for coordinated direction of
-gaze. Studying human eye movement has significant implications for improving our
-understanding of the oculomotor system and treating visuomotor disorders. Over
-the past decades, biomechanics simulation has provided the means to analyze
-different human movements. The same principles can be used to analyze visual
-tasks by modeling the musculoskeletal properties of the oculomotor
+gaze~\cite{Lee2006}. Studying human eye movement has significant implications
+for improving our understanding of the oculomotor system and treating visuomotor
+disorders. Over the past decades, biomechanics simulation has provided the means
+to analyze different human movements. The same principles can be used to analyze
+visual tasks by modeling the musculoskeletal properties of the oculomotor
 system. Consequently, this model can be used to investigate muscle activation
@@
 trials have provided a profound knowledge of how the \gls{em} act on the eye
-globe, the resistive tension of the surrounding tissues and the length-tension
-relationship of the muscles. Various computational models of the extraocular
-muscles and orbital mechanics have been proposed, which provide insight and
-scientific bases for oculomotor biomechanics, control of eye movement and
-binocular misalignment. These models focus on the realism of muscle behavior and
-they were based on the viscoelastic properties and physiological data \gls{em}.
+globe~\cite{Robinson1969a}, the resistive tension of the surrounding tissues and
+the length-tension relationship of the muscles~\cite{Collins1981}. Various
+computational models of the extraocular muscles and orbital mechanics have been
+proposed, which provide insight and scientific bases for oculomotor
+biomechanics, control of eye movement and binocular misalignment. These models
+focus on the realism of muscle behavior and they were based on the viscoelastic
+properties and physiological data \gls{em}.
 
@@
   Millard2013}. Such models, having the advantage of supporting dynamics
-simulation, are used in conjunction with brain level controllers \hl{seb}.
+simulation, are used in conjunction with brain level
+controllers~\cite{James2018}.
 
@@
 the constants and $\vec{\dot{q}} \inr{3}$ the rotational coordinates of the
-model. These forces serve the eye’s stabilization, and are modeled using
-\texttt{OpenSim}'s expression based coordinate force.
+model~\cite{Collins1981}. These forces serve the eye’s stabilization, and are
+modeled using \texttt{OpenSim}'s expression based coordinate force.
 
@@
 proposed computational eye model. In this study, we didn't attempt to model the
-muscle pulleys, which vary as a function of the model coordinates. Furthermore,
-the users should consider validation of the eye model based on the requirements
-of the targeted utility of the model and the variables of interests.
+muscle pulleys~\cite{Kono2002a}, which vary as a function of the model
+coordinates. Furthermore, the users should consider validation of the eye model
+based on the requirements of the targeted utility of the model and the variables
+of interests.
 
@@
 
-\bibliography*{mylibrary}
+\bibliography{mylibrary}
 \addcontentsline{toc}{chapter}{Bibliography}

### 2018-07-10 doc/manuscript.tex
correct manuscript

@@
 \begin{abstract}
-  Studying human eye movement has significant implications for improving our
+  Studying human eye movements has significant implications for improving our
   understanding of the oculomotor system and treating various visuomotor
@@
     \url{https://simtk.org/projects/eye}}. The model is implemented in
-  \texttt{OpenSim}~\cite{Delp2007}, which is an open-source framework for
-  modeling and simulation of musculoskeletal systems. The calibration of the
-  model parameters is based on physiological measurements of the human
-  eye~\cite{Iskander2018}. The model incorporates an eye globe, orbital
-  suspension tissues and six \gls{em}. The excitation and activation patterns
-  for a variety of targets can be calculated using a closed-loop fixation
-  controller that drives the model to perform saccadic movements in a \gls{fd}
-  manner. The controller minimizes the error between the desired saccadic
-  trajectory and the predicted movement. Consequently, this model enables the
-  investigation muscle activation patterns during static fixation and analyze
-  the dynamics of various eye movements.
+  \texttt{OpenSim}, which is an open-source framework for modeling and
+  simulation of musculoskeletal systems. The calibration of the model parameters
+  is based on physiological measurements of the human eye. The model
+  incorporates an eye globe, orbital suspension tissues and six extraocular
+  muscles. The excitation and activation patterns for a variety of targets can
+  be calculated using the proposed closed-loop fixation controller that drives
+  the model to perform saccadic movements in a forward dynamics manner. The
+  controller minimizes the error between the desired saccadic trajectory and the
+  predicted movement. Consequently, this model enables the investigation muscle
+  activation patterns during static fixation and analyze the dynamics of various
+  eye movements.
 \end{abstract}
@@
 disorders. Over the past decades, biomechanics simulation has provided the means
-to analyze different human movements. The same principles can be used to analyze
-visual tasks by modeling the musculoskeletal properties of the oculomotor
-system. Consequently, this model can be used to investigate muscle activation
-patterns during static fixation, analyze the dynamics of various eye movements,
-calculate metabolic costs and simulate different eye disorders, such as
+to analyze different human movements~\cite{Delp2007}. The same principles can be
+used to analyze visual tasks by modeling the musculoskeletal properties of the
+oculomotor system. Consequently, this model can be used to investigate muscle
+activation patterns during static fixation, analyze the dynamics of various eye
+movements, calculate metabolic costs and simulate eye disorders, such as
 different forms of strabismus. Furthermore, it can be easily integrated with
@@
 
-Eye movements are a generated from a coordination of the six \gls{em}. Clinical
-trials have provided a profound knowledge of how the \gls{em} act on the eye
-globe~\cite{Robinson1969a}, the resistive tension of the surrounding tissues and
-the length-tension relationship of the muscles~\cite{Collins1981}. Various
+Saccadic movements are a generated from a coordination of the six
+\gls{em}. Clinical trials have provided a profound knowledge on the properties
+of the \gls{em} and their line of action on the eye globe~\cite{Robinson1969a},
+the resistive tension of the surrounding tissues~\cite{Collins1981} and the
+length-tension relationship of the muscles~\cite{Iskander2018}. Various
 computational models of the extraocular muscles and orbital mechanics have been
@@
 The first 3D biomechanical model was developed by~\cite{Robinson1964a,
-  Robinson1969}, who simplified the formulation by considering the elasticity of
-the \gls{em} ignoring their dynamics. The model incorporates anatomically
-realistic muscle paths and empirical innervation-length-tension
+  Robinson1969}, who simplified the formulation by only considering the
+elasticity of the \gls{em} ignoring their dynamics. The model incorporates
+anatomically realistic muscle paths and empirical innervation-length-tension
 relationships. To study the neural control of rapid saccadic movements, models
@@
 by a solid sphere with 0.012 m radius. The eyeball was constructed in
-\texttt{Blender}, an open-source software 3D creation software. We used a
-spherical mesh with 32 segments and 12 rings, to construct the vitreous humor
-(body) as solid sphere and a conical plate to construct the cornea. The weight
-of an average human eye is 0.0075 kg and the moment of inertia can be calculated
-similarly as in the case of a spherical homogeneous and isotropic object with
-radius 0.012 m ($\mat{I} = 2/5 m r^2$ at the center of mass).
+\texttt{Blender}, using a spherical mesh with 32 segments and 12 rings, to
+construct the vitreous humor (body) as solid sphere and a conical plate to
+construct the cornea. The weight of an average human eye is 0.0075 kg and the
+moment of inertia can be calculated similarly as in the case of a spherical
+homogeneous and isotropic object with ($\mat{I} = 2/5 m r^2$ at the center of
+mass).
 
@@
 the image of the object. The \gls{lr} and \gls{mr} muscles form an antagonistic
-pair to produce horizontal eye movements. The \gls{sr} and \gls{ir} muscles form
-the vertical antagonist pair, which mainly controls vertical eye movement and
-also affects rotation about the horizontal plane and the line of sight
-(secondary action) due to insertion positions and the path of the muscles. The
-\gls{so} muscle passes through the cartilaginous trochlea attached to the
-orbital wall, which reflects the \gls{so} path by 51 deg. The \gls{io} muscle
-originates from the orbital wall anteroinferior to the globe center and inserts
-on the sclera posterior to the globe equator. The primary actions of \gls{so}
-and \gls{io} cause rotation of the globe around the visual axis and vertical
-movement.
+pair that produce horizontal eye movements. The \gls{sr} and \gls{ir} muscles
+form the vertical antagonist pair, which mainly controls vertical eye movement
+and also affects rotation about the horizontal plane and the line of sight
+(secondary action) due to insertion positions. The \gls{so} muscle passes
+through the cartilaginous trochlea attached to the orbital wall, which reflects
+the \gls{so} path by 51 deg. The \gls{io} muscle originates from the orbital
+wall anteroinferior to the globe center and inserts on the sclera posterior to
+the globe equator. The primary actions of \gls{so} and \gls{io} cause rotation
+of the globe around the visual axis and vertical movement.
 
@@
 physiological measurements~\cite{Iskander2018}, with some minor adaptation so as
-to prevent penetration into the eye globe. Since no position was documented for
-the origin of the \gls{so}, a point close to the origins of the rectus muscles
-was chosen to match the fiber length in the primary position of the \gls{so}
-muscle.
+to prevent unrealistic penetration into the eye globe. Since no position was
+documented for the origin of the \gls{so}, a point close to the origins of the
+rectus muscles was chosen to match the fiber length in the primary position of
+the \gls{so} muscle.
 
 \begin{table}[ht]
-  \caption{Muscle path for the six \gls{em} (dimensions are given in
+  \caption{Muscle path points for the six \gls{em} (dimensions are given in
     meters).}\label{tab:muscle-path}
@@
 The Millard muscle model~\cite{Millard2013} has been adopted for the modeling of
-the \gls{em}, which allows to manually fit the reported data for the \gls{fl}
-curves. The muscles were modeled using the rigid tendon assumption that ignores
-the elasticity of the tendon. This means that the series element of the muscle
-model is not included (the tendon length $l^T$ is equal to the tendon slack
-length $l_s^T$). This assumption is valid when the ratio of the tendon length to
-the muscle length is less or equal to one, as the in the case of all
-\gls{em}. \gls{em} are considered parallel-fibered muscles, so the pennation
-angle is zero ($\alpha = 0$). Maximum isometric force $f_o^M$, optimal fiber
-length $l_o^M$ and tendon length $l^T$ are presented
+the \gls{em}, which permits parameterization of the \gls{fl} curves according to
+the experimental measured data. The muscles were modeled using the rigid tendon
+assumption that ignores the elasticity of the tendon. This means that the series
+element of the muscle model is not included (the tendon length $l^T$ is equal to
+the tendon slack length $l_s^T$). This assumption is valid when the ratio of the
+tendon length to the muscle length is less or equal to one. \gls{em} are
+considered parallel-fibered muscles, so the pennation angle is zero
+($\alpha = 0$). The values for the maximum isometric force $f_o^M$, optimal
+fiber length $l_o^M$ and tendon length $l^T$ are presented
 in~\autoref{tab:muscle-path}.
@@
 skeletal muscle. As shown in \autoref{fig:millard-curves}, we can fine-tune
-these curves so as to fit the experimental data available for the gls{lr}
-muscle. The following values where used for the active \gls{fl} curve:
+these curves so as to fit the experimental data available for the \gls{lr}
+muscle. The following values where used for the active
+(\autoref{tab:active-fl-curve}) and passive \gls{fl} curves.
 
-\begin{itemize}
-\item min norm active fiber length: 0.55
-\item transition norm fiver length: 0.7
-\item max norm active fiver length: 1.8
-\item shallow ascending slope: 2.4
-\item minimum value: 0.0
-\end{itemize}
-%
-and for the passive \gls{fl} curve accordingly:
-
-\begin{itemize}
-\item strain at zero force: -0.18
-\item strain at one norm force: 0.4
-\end{itemize}
+\begin{table}
+  \parbox{.45\linewidth}{
+    \centering
+    \caption{Parameters of the active \gls{fl} curve of the Millard muscle
+      model.}\label{tab:active-fl-curve}
+    \begin{tabular}{@{}cccccccccc@{}}
+      \toprule
+      \textbf{Parameter} & \textbf{Value} \\
+      \midrule
+      min norm active fiber length & 0.55 \\
+      transition norm fiver length & 0.7 \\
+      max norm active fiver length & 1.8 \\
+      shallow ascending slope &  2.4 \\
+      minimum value & 0.0 \\
+      \bottomrule
+    \end{tabular}
+  }
+  \quad
+  \parbox{.45\linewidth}{
+    \vspace{-1.5cm}
+    \centering
+    \caption{Parameters of the active \gls{fl} curve of the Millard muscle
+      model.}\label{tab:passive-fl-curve}
+    \begin{tabular}{@{}cccccccccc@{}}
+      \toprule
+      \textbf{Parameter} & \textbf{Value} \\
+      \midrule
+      strain at zero force & -0.18 \\
+      strain at one norm force & 0.4 \\
+      \bottomrule
+    \end{tabular}
+  }
+\end{table}
 
@@
     keepaspectratio]{passive-force-length-curve.png}\label{fig:passive-force-length-curve}}
-  \caption{The active and passive \gls{fl} curve definition of the Millard
+  \caption{The active and passive \gls{fl} curve definition for the Millard
     muscle model as implemented in \texttt{OpenSim}.}\label{fig:millard-curves}
@@
 testing the fidelity of the model. Due to lack of data describing the other
-muscles’ \gls{fl} relationships, we used the parameters found describing the
-normalized curves of active and passive \gls{fl} relationships of the gls{lr}
-for the other \gls{em} as well.
+muscles, we used the parameters describing the normalized curves of active and
+passive \gls{fl} relationships of the \gls{lr} for the rest of the \gls{em}.
 
@@
 maximum contraction velocity $v^{\text{max}}$. The maximum muscle contraction
-velocity is tuned so as to match the peak velocity of saccadic eye movement 15.7
-rad / s. Following this definition, the maximum muscle contraction velocity is
-given in optimal fiber length per seconds and it is thus different for each
-\gls{em}, as their optimal fiber length is different. Furthermore, because of
-the different structure of neural control of the eyes, activation and
-deactivation delays ($\tau_d = 5$ ms) are lower than in skeletal
-muscles. Finally, two separate wrapping spheres for the rectus muscles and the
-oblique muscles were created, to avoid abnormal changes on the \gls{fl} curve as
-the eyeball rotates in the three directions.
+velocity is tuned so as to match the peak velocity of saccadic eye movement
+$\omega^{\text{max}} = 15.7$ rad / s (900 deg / s). Following this definition,
+the maximum muscle contraction velocity is given in optimal fiber length per
+seconds and it is thus different for each \gls{em}, as their optimal fiber
+length is different ($v^{\text{max}} = \omega^{\text{max}} r /
+l_o^M$). Furthermore, because of the different structure of neural control of
+the eyes, activation and deactivation delays ($\tau_d = 5$ ms) are lower than in
+skeletal muscles. Finally, two separate wrapping spheres for the rectus muscles
+and the oblique muscles were created, to avoid abnormal changes on the \gls{fl}
+curve as the eyeball rotates.
 
@@
 
-The passive connective tissues of the eyeball apply a restoring force, which
-brings the globe back to the central position when the net force from the
-\gls{em} is zero. These tissues include all non-muscular suspensory tissues,
-such as Tenon’s capsule, the optic nerve, the fat pad and the conjunctiva. The
-force-displacement curve of the net elasticity can be represented as
+The passive connective tissues apply a restoring force, which brings the globe
+back to the central position when the net force from the \gls{em} is zero. These
+tissues include all non-muscular suspensory tissues, such as Tenon’s capsule,
+the optic nerve, the fat pad and the conjunctiva. The force-displacement curve
+of the net elasticity can be represented as
 
@@
 % 
-where, $\vec{f}_t$ represents the passive tissue forces, $k_p= 0.002225$ N m /
-rad, $k_c= 34.5297$ N m / (rad\textsuperscript{3}) and $k_v= 0.002$ N m s / rad
-the constants and $\vec{\dot{q}} \inr{3}$ the rotational coordinates of the
-model~\cite{Collins1981}. These forces serve the eye’s stabilization, and are
-modeled using \texttt{OpenSim}'s expression based coordinate force.
+where, $\vec{f}_t$ is the passive tissue forces, $k_p= 0.002225$ N m / rad,
+$k_c= 34.5297$ N m / (rad\textsuperscript{3}) and $k_v= 0.002$ N m s / rad the
+constants and $\vec{\dot{q}} \inr{3}$ the rotational coordinates of the
+model~\cite{Collins1981}. These forces are modeled using \texttt{OpenSim}'s
+expression based coordinate force.
 
@@
 parameters of the controller are: the desired horizontal $\theta_H$ and vertical
-$\theta_V$ fixation angles (in degrees), the saccade onset and velocity, and the
-gains of \gls{pd} tracking controller ($k_p$, $k_d $). A sigmoid function is
-used for generating smooth saccade trajectories
+$\theta_V$ fixation angles (in degrees), the saccade onset and velocity (in
+degrees), and the gains of \gls{pd} tracking controller ($k_p$, $k_d $). A
+sigmoid function is used for generating smooth saccade trajectories
 
@@
 and velocity at time $t$, $a$ the magnitude of the trajectory, $b$ the slope and
-$t_0$ a time shift constant. Provided a fixation goal $\theta_g$ and a desired
-saccade velocity $\dot{\theta}_g$ the parameters of the sigmoid function are
-defined as follows $a = \theta_g$, $b = 2 \dot{\theta}_g / \theta_g$. The
-\gls{pd} tracking controller has the following form
+$t_0$ a time shift constant. Provided a fixation goal $\theta_g$, a desired
+saccade velocity $\dot{\theta}_g$ and a saccade onset $t_g$ the parameters of
+the sigmoid function are defined as follows $a = \theta_g$,
+$b = 2 \dot{\theta}_g / \theta_g$ and $t_0 = t_s$. The \gls{pd} tracking
+controller has the following form
 
@@
 The sign and magnitude of $\ddot{\theta}(t)$ both for the horizontal and
-vertical coordinates of the fixation target is used to actuate the corresponding
-muscles in order to achieve the desired goal. \autoref{fig:model} presents an
-instance of the model during simulation with the corresponding muscles
-activated. \autoref{fig:simulated-saccade} depicts the simulated coordinates,
-speeds and estimated \gls{em} activations that reproduce the desired saccade
-trajectory.
+vertical displacement of the fixation target is used to actuate the
+corresponding muscles in order to achieve the desired goal. \autoref{fig:model}
+presents an instance of the model during simulation with the corresponding
+muscles activated. \autoref{fig:simulated-saccade} depicts the simulated
+coordinates, speeds and estimated \gls{em} activations that reproduce the
+desired saccade trajectory.
 
@@
   \caption{Model with a fixation at $\theta_H = -15$ deg $\theta_V = 15$ deg
-    during simulation. Blue denotes low and red high activation
-    values.}\label{fig:model}
+    during simulation. Blue denotes low and red high muscle activation
+    levels.}\label{fig:model}
 \end{figure}
@@
 improvement, which will enhance the accuracy and the predictability of the
-proposed computational eye model. In this study, we didn't attempt to model the
+proposed computational model. In this study, we didn't attempt to model the
 muscle pulleys~\cite{Kono2002a}, which vary as a function of the model
-coordinates. Furthermore, the users should consider validation of the eye model
-based on the requirements of the targeted utility of the model and the variables
-of interests.
+coordinates. Therefore, the users should consider performing further validation
+of the eye model based on the requirements of the targeted utility and the
+variables of interests.
 
@@
 \bibliography{mylibrary}
-\addcontentsline{toc}{chapter}{Bibliography}
+% \addcontentsline{toc}{chapter}{Bibliography}
 

### 2018-07-16 doc/manuscript.tex
centering
will start working on simulation section

@@
 \usepackage{cancel}
+\usepackage{siunitx}            % include before arydshln
 \usepackage{arydshln}         % matrix dash lines
@@
 \newacronym{fv}{F-V}{Force-Velocity}
+\newacronym{fpe}{F-PE}{Passive-Force-Length}
+\newacronym{ft}{F-T}{Tendon-Force-Length}
 \makeglossaries{}
@@
   disorders. An open-source musculoskeletal model of the human eye, that can be
-  used for kinematics and dynamics analysis, is implemented based on the data
+  used for kinematics and dynamics analysis, is constructed based on the data
   reported in literature and made publicly available\footnote{SimTK project:
@@
 
-The orbital plant consists of the globe (eyeball), three pairs of extraocular
-muscles and the connective passive tissues. The size of an emmetropic human
-adult eye is approximately 0.0242 m (transverse, horizontal), 0.0237 m
-(sagittal, vertical), 0.022–0.0248 m (axial, anteroposterior) with no
+The eye model consists of the eyeball, three pairs of \gls{em} and the
+connective passive tissues. The size of an emmetropic human adult eye is
+approximately $0.0242 \si{\m}$ (transverse, horizontal), $0.0237 \si{\m}$
+(sagittal, vertical), $0.022 - 0.0248 \si{\m}$ (axial, anteroposterior) with no
 significant difference between sexes and age groups. In the transverse diameter,
-the eyeball size may vary from 0.021 m to 0.027 m. Thus, it can be approximated
-by a solid sphere with 0.012 m radius. The eyeball was constructed in
-\texttt{Blender}, using a spherical mesh with 32 segments and 12 rings, to
-construct the vitreous humor (body) as solid sphere and a conical plate to
-construct the cornea. The weight of an average human eye is 0.0075 kg and the
-moment of inertia can be calculated similarly as in the case of a spherical
-homogeneous and isotropic object with ($\mat{I} = 2/5 m r^2$ at the center of
-mass).
+the eyeball size may vary from $0.021 \si{\m}$ to $0.027 \si{\m}$, thus it can
+be approximated by a solid sphere of radius $r = 0.012 \si{\m}$. The eyeball was
+constructed in \texttt{Blender}, using a spherical mesh with 32 segments and 12
+rings, to construct the vitreous humor (body) as solid sphere and a conical
+plate to construct the cornea.The weight of an average human eye is
+$m = 0.0075 \si{\kg}$ and the moment of inertia can be calculated assuming a
+spherical homogeneous and isotropic model $I = 2 / 5 m r^2$. The eye has three
+rotational \gls{dofs} with respect to the frame of reference, namely
+incyclotosion-excyclotosion ($x$-axis), adduction-abduction ($y$-axis) and
+supraduction-infraduction ($z$-axis).
 
@@
 
-The model relies on the passive pulley assumption in order to keep it simple and
-provide faster simulation speed. \autoref{tab:muscle-path} shows the positions
-of muscle pulleys, as well as the origin and insertion points of the \gls{em},
-defined in the local body coordinates of the globe. The data is based on
-physiological measurements~\cite{Iskander2018}, with some minor adaptation so as
-to prevent unrealistic penetration into the eye globe. Since no position was
-documented for the origin of the \gls{so}, a point close to the origins of the
-rectus muscles was chosen to match the fiber length in the primary position of
-the \gls{so} muscle.
+The model relies on the passive pulley assumption. \autoref{tab:muscle-path}
+shows the positions of the origin, insertion and passive muscle pulleys for the
+\gls{em}, defined in the local body coordinates of the eyeball. The data are
+based on physiological measurements~\cite{Iskander2018}, with some minor
+modification so as to prevent unrealistic muscle-surface penetration. Since no
+position was documented for the origin of the \gls{so}, a point close to the
+origins of the rectus muscles was chosen to match the fiber length in the
+primary position of the \gls{so} muscle.
 
 \begin{table}[ht]
-  \caption{Muscle path points for the six \gls{em} (dimensions are given in
-    meters).}\label{tab:muscle-path}
+  \centering
+  \caption{Muscle path points for the six \gls{em} defined in the local frame of
+    the eyeball (dimensions are given in meters).}\label{tab:muscle-path}
   \begin{tabular}{@{}cccccccccc@{}}
@@
 The Millard muscle model~\cite{Millard2013} has been adopted for the modeling of
-the \gls{em}, which permits parameterization of the \gls{fl} curves according to
-the experimental measured data. The muscles were modeled using the rigid tendon
-assumption that ignores the elasticity of the tendon. This means that the series
-element of the muscle model is not included (the tendon length $l^T$ is equal to
-the tendon slack length $l_s^T$). This assumption is valid when the ratio of the
-tendon length to the muscle length is less or equal to one. \gls{em} are
-considered parallel-fibered muscles, so the pennation angle is zero
-($\alpha = 0$). The values for the maximum isometric force $f_o^M$, optimal
-fiber length $l_o^M$ and tendon length $l^T$ are presented
-in~\autoref{tab:muscle-path}.
+the \gls{em}, permitting parameterization of the characteristic curves according
+to the experimental measured data. The muscles were modeled using the rigid
+tendon assumption that ignores the elasticity of the tendon. This means that the
+series element of the muscle model is not included (the tendon length $l^T$ is
+equal to the tendon slack length $l_s^T$). \gls{em} are considered
+parallel-fibered muscles, so the pennation angle is zero ($\alpha = 0$). The
+values for the maximum isometric force $f_o^M$, optimal fiber length $l_o^M$ and
+tendon length $l^T$ are presented in~\autoref{tab:muscle-parameters}.
 
-The active and passive \gls{fl} curves for the \gls{em} differ from that of a
-skeletal muscle. As shown in \autoref{fig:millard-curves}, we can fine-tune
-these curves so as to fit the experimental data available for the \gls{lr}
-muscle. The following values where used for the active
-(\autoref{tab:active-fl-curve}) and passive \gls{fl} curves.
+The \gls{fl} and \gls{fpe} characteristic curves of the \gls{em} differ
+significantly from those of a skeletal muscle. As shown in
+\autoref{fig:millard-curves}, we can fine-tune the curve parameters so as to fit
+the experimental data available for the \gls{lr} muscle. The values for the
+\gls{fl} and \gls{fpe} characteristic curves are summarized in
+Tables~\ref{tab:fl-curve} and~\ref{tab:fpe-curve}, respectively. Due to lack of
+data describing the other muscles, the parameters found for the \gls{lr} muscle
+where used for the rest of the \gls{em}.
 
 \begin{table}
+  \centering
   \parbox{.45\linewidth}{
     \centering
-    \caption{Parameters of the active \gls{fl} curve of the Millard muscle
-      model.}\label{tab:active-fl-curve}
+    \caption{Parameters of the \gls{fl} characteristic curve for the
+      \gls{em}.}\label{tab:fl-curve}
     \begin{tabular}{@{}cccccccccc@{}}
@@
     \centering
-    \caption{Parameters of the active \gls{fl} curve of the Millard muscle
-      model.}\label{tab:passive-fl-curve}
+    \caption{Parameters of the \gls{fpe} characteristic curve for the for the
+      \gls{em}.}\label{tab:fpe-curve}
     \begin{tabular}{@{}cccccccccc@{}}
@@
 \begin{figure}[ht]
-  \subfloat[Active \gls{fl} curve]{\includegraphics[width=0.5\textwidth,
+  \centering
+  \subfloat[\gls{fl} curve]{\includegraphics[width=0.5\textwidth,
     keepaspectratio]{active-force-length-curve.png}\label{fig:active-force-length-curve}}
-  \subfloat[Passive \gls{fl} curve]{\includegraphics[width=0.5\textwidth,
+  \subfloat[\gls{fl} curve]{\includegraphics[width=0.5\textwidth,
     keepaspectratio]{passive-force-length-curve.png}\label{fig:passive-force-length-curve}}
-  \caption{The active and passive \gls{fl} curve definition for the Millard
-    muscle model as implemented in \texttt{OpenSim}.}\label{fig:millard-curves}
+  \caption{The \gls{fl} and \gls{fpe} curve definition for the Millard muscle
+    model as implemented in \texttt{OpenSim}.}\label{fig:millard-curves}
 \end{figure}
-
-The parameters that describe the above relationships were chosen to fit the
-curves reported in~\cite{Iskander2018}, by matching the \gls{fl} relationship at
-maximum activation of the \gls{lr} muscle. This represents the first part of
-testing the fidelity of the model. Due to lack of data describing the other
-muscles, we used the parameters describing the normalized curves of active and
-passive \gls{fl} relationships of the \gls{lr} for the rest of the \gls{em}.
 
@@
 muscles. Despite that, the default Millard \gls{fv} curve was used for the six
-\gls{em}, since the behavior of the selected muscle model depends mainly on the
+\gls{em}, since the behavior of the selected muscle model depends heavily on the
 maximum contraction velocity $v^{\text{max}}$. The maximum muscle contraction
 velocity is tuned so as to match the peak velocity of saccadic eye movement
-$\omega^{\text{max}} = 15.7$ rad / s (900 deg / s). Following this definition,
-the maximum muscle contraction velocity is given in optimal fiber length per
-seconds and it is thus different for each \gls{em}, as their optimal fiber
-length is different ($v^{\text{max}} = \omega^{\text{max}} r /
-l_o^M$). Furthermore, because of the different structure of neural control of
-the eyes, activation and deactivation delays ($\tau_d = 5$ ms) are lower than in
-skeletal muscles. Finally, two separate wrapping spheres for the rectus muscles
-and the oblique muscles were created, to avoid abnormal changes on the \gls{fl}
-curve as the eyeball rotates.
+$\omega^{\text{max}} = 15.7 \si{\radian / \second}$
+($900 \si{\degree / \second}$). Following this definition, the maximum muscle
+contraction velocity is given in optimal fiber length per seconds and it is thus
+different for each \gls{em}, as their optimal fiber length is different
+($v^{\text{max}} = \omega^{\text{max}} r / l_o^M$). Furthermore, since the optic
+nerve is much shorter that the average muscle nerve, activation and deactivation
+delays ($\tau_a = \tau_d = 5 \si{\milli\second}$) are smaller. Finally, two
+separate wrapping spheres for the rectus muscles and the oblique muscles were
+created, to avoid abnormal changes on the \gls{fl} curve as the eyeball rotates.
 
 \begin{table}[ht]
+  \centering
   \caption{Millard muscle parameters for the
@@
     \toprule
-    \thead{Muscle}
-    & \thead{Maximum Isometric \\ Force (N)}
+    \thead{Muscle \\ \quad}
+    & \thead{Maximum Isometric \\ Force (N)}  % math \si not in bold
     & \thead{Optimal Fiber \\ Length (m)}
@@
 
-The passive connective tissues apply a restoring force, which brings the globe
+The passive connective tissues apply a restoring force, which brings the eyeball
 back to the central position when the net force from the \gls{em} is zero. These
-tissues include all non-muscular suspensory tissues, such as Tenon’s capsule,
+tissues include all non-muscular suspensory tissues, such as Tenon's capsule,
 the optic nerve, the fat pad and the conjunctiva. The force-displacement curve
@@
 %
-where, $\vec{f}_t$ is the passive tissue forces, $k_p= 0.002225$ N m / rad,
-$k_c= 34.5297$ N m / (rad\textsuperscript{3}) and $k_v= 0.002$ N m s / rad the
-constants and $\vec{\dot{q}} \inr{3}$ the rotational coordinates of the
-model~\cite{Collins1981}. These forces are modeled using \texttt{OpenSim}'s
-expression based coordinate force.
+where, $\vec{f}_t$ is the passive tissue forces,
+$k_p= 0.002225 \si{\N \m / \radian}$, $k_c= 34.5297 \si{\N \m / \radian^3}$ and
+$k_v= 0.002$ $\si{\N \m / (\radian / \s)}$ the constants and
+$\vec{\dot{q}} \inr{3}$ the rotational coordinates of the
+model~\cite{Collins1981} \hl{TODO}. These forces are modeled using
+\texttt{OpenSim}'s expression based coordinate force.
 
@@
   \centering
-  \includegraphics[width=.6\textwidth]{model.png}
+  \includegraphics[width=.6\textwidth]{eye-model.png}
   \caption{Model with a fixation at $\theta_H = -15$ deg $\theta_V = 15$ deg

### 2018-07-16 source_code/src/fixation_controller_plugin/FixationController.cpp
update results

@@ -32,7 +32,7 @@ FixationController::FixationController() : Controller() {
 }
 
 void FixationController::constructProperties() {
-    constructProperty_thetaH(0);
+    constructProperty_thetaH(-15);
     constructProperty_thetaV(0);
     constructProperty_kpH(50);
     constructProperty_kdH(1.5);
@@ -41,7 +41,7 @@ void FixationController::constructProperties() {
     constructProperty_kpT(100);
     constructProperty_kdT(0.5);
     constructProperty_saccade_onset(0.5);
-    constructProperty_saccade_velocity(100);
+    constructProperty_saccade_velocity(600);
 }
 
 void FixationController::computeControls(const State& s, Vector& controls) const {
@@ -67,19 +67,22 @@ void FixationController::computeControls(const State& s, Vector& controls) const
     double xdes = 0;
     double xdesv = 0;
     double xdesa = 0;
+
     auto sigmY = sigmoid(t,
                          get_saccade_onset(),
                          thetaH,
                          2 * desSaccadeVelocity / abs(thetaH));
-    double ydes = 0 + sigmY[0];
-    double ydesv = sigmY[1];
+
+    double ydes = isnan(sigmY[0]) ? 0 : sigmY[0];
+    double ydesv = isnan(sigmY[1]) ? 0 : sigmY[1];
     double ydesa = sigmY[2];
+
     auto sigmZ = sigmoid(t,
                          get_saccade_onset(),
                          thetaV,
                          2 * desSaccadeVelocity / abs(thetaV));
-    double zdes = 0 + sigmZ[0];
-    double zdesv = sigmZ[1];
+    double zdes = isnan(sigmZ[0]) ? 0 : sigmZ[0];
+    double zdesv = isnan(sigmZ[1]) ? 0 : sigmZ[1];
     double zdesa = sigmZ[2];
 
     // Get the current position and velocity
@@ -88,6 +91,7 @@ void FixationController::computeControls(const State& s, Vector& controls) const
     double z = zCoord.getValue(s); double zv = zCoord.getSpeedValue(s);
 
     // The sum of errors are used as Excitation levels in the model
+    // acceleration is unused because there are problems with the sigmoid
     double sumErrX = 0 * xdesa + get_kpT() * (xdes - x) + get_kdT() * (xdesv - xv);
[…]

### 2018-07-16 source_code/src/include/Settings.h

@@ -3,6 +3,6 @@
 
 #include <string>
 
-std::string MODEL_DIR = "C:/Users/mitkof6/Desktop/upat_eye_model/source_code/../model/";
+std::string MODEL_DIR = "D:/Desktop/upat_eye_model/source_code/../model/";
 
 #endif

### 2018-07-16 source_code/src/simulation/Simulation.cpp

@@ -43,19 +43,19 @@ void addExpressionCoordinateForce(Model* model) {
     // recent versions may not have this problem.
     auto add_adb_tissue =
         new ExpressionBasedCoordinateForce("r_eye_add_abd",
-                                           "-0.002225*q-34.5297*0.0001*q^3-0.002*qdot");
+                                           "-0.002225*q-34.5297*0.0001*q^3-0*0.002*qdot");
     add_adb_tissue->setName("add_adb_tissue");
     model->addForce(add_adb_tissue);
 
     auto sup_inf_tissue =
         new ExpressionBasedCoordinateForce("r_eye_sup_inf",
-                                           "-0.002225*q-34.5297*0.0001*q^3-0.002*qdot");
+                                           "-0.002225*q-34.5297*0.0001*q^3-0*0.002*qdot");
     sup_inf_tissue->setName("sup_inf_tissue");
     model->addForce(sup_inf_tissue);
 
     auto inc_exc_tissue =
         new ExpressionBasedCoordinateForce("r_eye_inc_exc",
-                                           "-0.002225*q-34.5297*0.0001*q^3-0.002*qdot");
+                                           "-0.002225*q-34.5297*0.0001*q^3-0*0.002*qdot");
     inc_exc_tissue->setName("inc_exc_tissue");
     model->addForce(inc_exc_tissue);
 }
@@ -79,7 +79,7 @@ void simulateModel() {
     controller->set_kpT(100);
     controller->set_kdT(0.5);
     controller->set_saccade_onset(0.5);
-    controller->set_saccade_velocity(100); // deg / s
+    controller->set_saccade_velocity(600); // deg / s
     model.addController(controller);
 
     // Build and initialize model

### 2018-07-16 source_code/src/fixation_controller_plugin/FixationController.cpp
-1515

@@ -13,17 +13,17 @@ using namespace SimTK;
  * ----------
  * t : time to be evaluated
  * t0 : delay
- * A : magnitude
- * B : slope
+ * a : magnitude
+ * b : slope
  *
  * Returns
  * -------
  * (y, y', y'')
  */
-Vec3 sigmoid(double t, double t0, double A, double B) {
-    return Vec3(A * (tanh(B * (t - t0)) + 1) / 2,
-                -(A * B * (pow(tanh(B * (t - t0)), 2) - 1)) / 2,
-                A * pow(B, 2) * tanh(B * (t - t0)) * (pow(tanh(B * (t - t0)), 2) - 1)
+Vec3 sigmoid(double t, double t0, double a, double b) {
+    return Vec3(a * (tanh(b * (t - t0)) + 1) / 2,
+                -(a * b * (pow(tanh(b * (t - t0)), 2) - 1)) / 2,
+                a * pow(b, 2) * tanh(b * (t - t0)) * (pow(tanh(b * (t - t0)), 2) - 1)
     );
 }
 

### 2018-07-16 doc/manuscript.tex
prepare manuscript

@@
 \DeclareGraphicsExtensions{.pdf,.png,.jpg}
+\usepackage{wrapfig}
 \usepackage{caption}
@@
 \begin{equation}\label{equ:passive-tissue}
-  \vec{f}_t = -k_p \vec{q} - k_c 10^{-3} \vec{q}^3 - k_d * \vec{\dot{q}}
+  \vec{f}_t = -k_p \vec{q} - k_c 10^{-3} \vec{q}^3 - k_v * \vec{\dot{q}}
 \end{equation}
@@
 $\vec{\dot{q}} \inr{3}$ the rotational coordinates of the
-model~\cite{Collins1981} \hl{TODO}. These forces are modeled using
-\texttt{OpenSim}'s expression based coordinate force.
+model~\cite{Collins1981}. These forces are modeled using \texttt{OpenSim}'s
+expression based coordinate force.
 
@@
   \centering
-  \includegraphics[width=.6\textwidth]{eye-model.png}
-  \caption{Model with a fixation at $\theta_H = -15$ deg $\theta_V = 15$ deg
-    during simulation. Blue denotes low and red high muscle activation
-    levels.}\label{fig:model}
+  \includegraphics[width=.7\textwidth]{eye-model.png}
+  \caption{Model with a fixation at $\theta_H = -15 \si{\degree}$,
+    $\theta_V = 15 \si{\degree}$ during simulation. Blue denotes low and red high
+    muscle activation levels.}\label{fig:model}
 \end{figure}
@@
 \begin{figure}[ht]
+  \vspace{-1.6cm}
   \centering
-  \includegraphics[width=1.\textwidth]{UPAT_Eye_Model_Passive_Pulleys_v3_States[-15][15].pdf}
-  \caption{Simulated saccade response with a fixation at $\theta_H = -15$ deg
-    $\theta_V = 15$ deg. The left subplot represents the simulated generalized
-    coordinates; the middle the coordinate speeds; right the estimated \gls{em}
+  \subfloat[$\theta_H = 15 \si{\degree}$, $\theta_V = 0 \si{\degree}$, $b = 600
+  \si{\degree / \s}$ and $k_v = 0 \si{\N \m / (\radian / \s)}$]{\includegraphics[width=1.\textwidth,
+    keepaspectratio]{UPAT_Eye_Model_Passive_Pulleys_v3_State_h15v0kv0.pdf}\label{fig:simulated-saccade-a}} 
+  \\
+  \subfloat[$\theta_H = 0 \si{\degree}$, $\theta_V = 15 \si{\degree}$, $b = 600
+  \si{\degree / \s}$ and $k_v = 0 \si{\N \m / (\radian / \s)}$]{\includegraphics[width=1.\textwidth,
+    keepaspectratio]{UPAT_Eye_Model_Passive_Pulleys_v3_Statesh0v15kv0.pdf}\label{fig:simulated-saccade-b}}
+  \\
+  \subfloat[$\theta_H = -15 \si{\degree}$, $\theta_V = 15 \si{\degree}$, $b = 600
+  \si{\degree / \s}$ and $k_v = 0 \si{\N \m / (\radian / \s)}$]{\includegraphics[width=1.\textwidth,
+    keepaspectratio]{UPAT_Eye_Model_Passive_Pulleys_v3_Statesh-15v15kv0.pdf}\label{fig:simulated-saccade-c}}
+  \\
+  \subfloat[$\theta_H = -15 \si{\degree}$, $\theta_V = 15 \si{\degree}$, $b = 100
+  \si{\degree / \s}$ and $k_v = 0.002 \si{\N \m / (\radian / \s)}$]{\includegraphics[width=1.\textwidth,
+    keepaspectratio]{UPAT_Eye_Model_Passive_Pulleys_v3_Statesh15v-15kv0002.pdf}\label{fig:simulated-saccade-d}}
+  \caption{Simulated saccade response for different model parameters. The left
+    subplot represents the simulated generalized coordinates; the middle the
+    coordinate speeds; right the estimated \gls{em}
     activations.}\label{fig:simulated-saccade}

### 2018-07-16 model/plot_results.py

@@ -90,93 +90,17 @@ def index_containing_substring(list_str, pattern):
     return indices
 
 
-def savitzky_golay(y, window_size, order, deriv=0, rate=1):
-    """Smooth (and optionally differentiate) data with a Savitzky-Golay filter.
-    The Savitzky-Golay filter removes high frequency noise from data.
-    It has the advantage of preserving the original shape and
-    features of the signal better than other types of filtering
-    approaches, such as moving averages techniques.
-    Parameters
-    ----------
-    y : array_like, shape (N,)
-        the values of the time history of the signal.
-    window_size : int
-        the length of the window. Must be an odd integer number.
-    order : int
-        the order of the polynomial used in the filtering.
-        Must be less then `window_size` - 1.
-    deriv: int
-        the order of the derivative to compute (default = 0 means only smoothing)
-    Returns
-    -------
-    ys : ndarray, shape (N)
-        the smoothed signal (or it's n-th derivative).
-    Notes
-    -----
-    The Savitzky-Golay is a type of low-pass filter, particularly
-    suited for smoothing noisy data. The main idea behind this
-    approach is to make for each point a least-square fit with a
-    polynomial of high order over a odd-sized window centered at
-    the point.
-    Examples
-    --------
-    t = np.linspace(-4, 4, 500)
-    y = np.exp( -t**2 ) + np.random.normal(0, 0.05, t.shape)
-    ysg = savitzky_golay(y, window_size=31, order=4)
-    import matplotlib.pyplot as plt
-    plt.plot(t, y, label='Noisy signal')
-    plt.plot(t, np.exp(-t**2), 'k', lw=1.5, label='Original signal')
-    plt.plot(t, ysg, 'r', label='Filtered signal')
-    plt.legend()
-    plt.show()
-    References
-    ----------
-    .. [1] A. Savitzky, M. J. E. Golay, Smoothing and Differentiation of
-       Data by Simplified Least Squares Procedures. Analytical
-       Chemistry, 1964, 36 (8), pp 1627-1639.
-    .. [2] Numerical Recipes 3rd Edition: The Art of Scientific Computing
-       W.H. Press, S.A. Teukolsky, W.T. Vetterling, B.P. Flannery
[…]

### 2018-07-16 source_code/src/fixation_controller_plugin/FixationController.cpp
figure

@@ -91,7 +91,7 @@ void FixationController::computeControls(const State& s, Vector& controls) const
     double z = zCoord.getValue(s); double zv = zCoord.getSpeedValue(s);
 
     // The sum of errors are used as Excitation levels in the model
-    // acceleration is unused because there are problems with the sigmoid
+    // acceleration is unused because causes problems
     double sumErrX = 0 * xdesa + get_kpT() * (xdes - x) + get_kdT() * (xdesv - xv);
     double sumErrY = 0 * ydesa + get_kpH() * (ydes - y) + get_kdH() * (ydesv - yv);
     double sumErrZ = 0 * zdesa + get_kpV() * (zdes - z) + get_kdV() * (zdesv - zv);

### 2018-07-16 source_code/src/simulation/Simulation.cpp

@@ -43,19 +43,19 @@ void addExpressionCoordinateForce(Model* model) {
     // recent versions may not have this problem.
     auto add_adb_tissue =
         new ExpressionBasedCoordinateForce("r_eye_add_abd",
-                                           "-0.002225*q-34.5297*0.0001*q^3-0*0.002*qdot");
+                                           "-0.002225*q-34.5297*0.0001*q^3-1*0.002*qdot");
     add_adb_tissue->setName("add_adb_tissue");
     model->addForce(add_adb_tissue);
 
     auto sup_inf_tissue =
         new ExpressionBasedCoordinateForce("r_eye_sup_inf",
-                                           "-0.002225*q-34.5297*0.0001*q^3-0*0.002*qdot");
+                                           "-0.002225*q-34.5297*0.0001*q^3-1*0.002*qdot");
     sup_inf_tissue->setName("sup_inf_tissue");
     model->addForce(sup_inf_tissue);
 
     auto inc_exc_tissue =
         new ExpressionBasedCoordinateForce("r_eye_inc_exc",
-                                           "-0.002225*q-34.5297*0.0001*q^3-0*0.002*qdot");
+                                           "-0.002225*q-34.5297*0.0001*q^3-1*0.002*qdot");
     inc_exc_tissue->setName("inc_exc_tissue");
     model->addForce(inc_exc_tissue);
 }
@@ -70,8 +70,8 @@ void simulateModel() {
     // Create a controller
     FixationController* controller = new FixationController();
     controller->setName("fixation_controller");
-    controller->set_thetaH(-15);
-    controller->set_thetaV(15);
+    controller->set_thetaH(15);
+    controller->set_thetaV(-15);
     controller->set_kpH(50);
     controller->set_kdH(1.5);
     controller->set_kpV(50);
@@ -79,7 +79,7 @@ void simulateModel() {
     controller->set_kpT(100);
     controller->set_kdT(0.5);
     controller->set_saccade_onset(0.5);
-    controller->set_saccade_velocity(600); // deg / s
+    controller->set_saccade_velocity(100); // deg / s
     model.addController(controller);
 
     // Build and initialize model

### 2018-07-16 doc/manuscript.tex
update results

@@
 coordinates, speeds and estimated \gls{em} activations that reproduce the
-desired saccade trajectory.
+desired saccade trajectory for different model parameters.
 

### 2018-07-16 model/plot_results.py

@@ -92,7 +92,7 @@ def index_containing_substring(list_str, pattern):
 
 def smooth(x):
     from scipy.signal import medfilt, gauss_spline
-    return medfilt(x, 9)
+    return medfilt(x, 1)
     # return x
 
 ###############################################################################
@@ -100,7 +100,8 @@ def smooth(x):
 
 # state = 'UPAT_Eye_Model_Passive_Pulleys_v3_State_h15v0kv0.sto'
 # state = 'UPAT_Eye_Model_Passive_Pulleys_v3_Statesh0v15kv0.sto'
-state = 'UPAT_Eye_Model_Passive_Pulleys_v3_Statesh-15v15kv0.sto'
+# state = 'UPAT_Eye_Model_Passive_Pulleys_v3_Statesh-15v15kv0.sto'
+state = 'UPAT_Eye_Model_Passive_Pulleys_v3_Statesh15v-15kv0002.sto'
 
 header, labels, data = readMotionFile(state)
 

### 2018-07-16 doc/manuscript.tex
minor update

@@
 
-The eye model consists of the eyeball, three pairs of \gls{em} and the
+The eye model consists of the eye globe, three pairs of \gls{em} and the
 connective passive tissues. The size of an emmetropic human adult eye is
@@
 significant difference between sexes and age groups. In the transverse diameter,
-the eyeball size may vary from $0.021 \si{\m}$ to $0.027 \si{\m}$, thus it can
-be approximated by a solid sphere of radius $r = 0.012 \si{\m}$. The eyeball was
-constructed in \texttt{Blender}, using a spherical mesh with 32 segments and 12
-rings, to construct the vitreous humor (body) as solid sphere and a conical
-plate to construct the cornea.The weight of an average human eye is
-$m = 0.0075 \si{\kg}$ and the moment of inertia can be calculated assuming a
-spherical homogeneous and isotropic model $I = 2 / 5 m r^2$. The eye has three
-rotational \gls{dofs} with respect to the frame of reference, namely
-incyclotosion-excyclotosion ($x$-axis), adduction-abduction ($y$-axis) and
-supraduction-infraduction ($z$-axis).
+the eye may vary from $0.021 \si{\m}$ to $0.027 \si{\m}$, thus it can be
+approximated by a solid sphere of radius $r = 0.012 \si{\m}$. The weight of an
+average human eye is $m = 0.0075 \si{\kg}$ and the moment of inertia can be
+calculated assuming a spherical homogeneous and isotropic model
+$I = 2 / 5 m r^2$. The eye has three rotational \gls{dofs}, namely
+incyclotosion-excyclotosion ($\vec{x}$-axis), adduction-abduction
+($\vec{y}$-axis) and supraduction-infraduction ($\vec{z}$-axis).
 
@@
 
-The model relies on the passive pulley assumption. \autoref{tab:muscle-path}
+The model relies on the passive pulley assumption. \autoref{tab:eye-muscle-path}
 shows the positions of the origin, insertion and passive muscle pulleys for the
-\gls{em}, defined in the local body coordinates of the eyeball. The data are
+\gls{em}, defined in the local body coordinates of the eye globe. The data are
 based on physiological measurements~\cite{Iskander2018}, with some minor
@@
   \caption{Muscle path points for the six \gls{em} defined in the local frame of
-    the eyeball (dimensions are given in meters).}\label{tab:muscle-path}
+    the eye globe (dimensions are given in meters).}\label{tab:eye-muscle-path}
   \begin{tabular}{@{}cccccccccc@{}}
@@
 values for the maximum isometric force $f_o^M$, optimal fiber length $l_o^M$ and
-tendon length $l^T$ are presented in~\autoref{tab:muscle-parameters}.
+tendon length $l^T$ are presented in~\autoref{tab:eye-muscle-parameters}.
+
+\begin{table}[ht]
+  \centering
+  \caption{Millard muscle parameters for the
+    \gls{em}.}\label{tab:eye-muscle-parameters}
+  \begin{tabular}{@{}cccccccccc@{}}
+    \toprule
+    \thead{Muscle \\ \quad}
+    & \thead{Maximum Isometric \\ Force (N)}  % math \si not in bold
+    & \thead{Optimal Fiber \\ Length (m)}
+    & \thead{Tendon Slack \\ Length (m)}
+    & \thead{Maximum Contraction \\ Velocity (m / s)} \\
+    \midrule
+    \gls{lr} & 1.4710 & 0.04898 & 0.0084 & 3.8483 \\
+    \gls{mr} & 1.5740 & 0.04084 & 0.0038 & 4.6155 \\
+    \gls{sr} & 1.1768 & 0.04487 & 0.0054 & 4.2009 \\
+    \gls{ir} & 1.4269 & 0.04549 & 0.0048 & 4.1437 \\
+    \gls{so} & 0.6031 & 0.03956 & 0.0265 & 4.7648 \\
+    \gls{io} & 0.5590 & 0.04110 & 0.0015 & 3.5863 \\
+    \bottomrule
+  \end{tabular}
+\end{table}
 
@@
 \gls{fl} and \gls{fpe} characteristic curves are summarized in
-Tables~\ref{tab:fl-curve} and~\ref{tab:fpe-curve}, respectively. Due to lack of
-data describing the other muscles, the parameters found for the \gls{lr} muscle
-where used for the rest of the \gls{em}.
+Tables~\ref{tab:eye-fl-curve} and~\ref{tab:eye-fpe-curve}, respectively.
 
@@
     \caption{Parameters of the \gls{fl} characteristic curve for the
-      \gls{em}.}\label{tab:fl-curve}
+      \gls{em}.}\label{tab:eye-fl-curve}
     \begin{tabular}{@{}cccccccccc@{}}
@@
     \centering
-    \caption{Parameters of the \gls{fpe} characteristic curve for the for the
-      \gls{em}.}\label{tab:fpe-curve}
+    \caption{Parameters of the \gls{fpe} characteristic curve for the
+      \gls{em}.}\label{tab:eye-fpe-curve}
     \begin{tabular}{@{}cccccccccc@{}}
@@
 separate wrapping spheres for the rectus muscles and the oblique muscles were
-created, to avoid abnormal changes on the \gls{fl} curve as the eyeball rotates.
-
-\begin{table}[ht]
-  \centering
-  \caption{Millard muscle parameters for the
-    \gls{em}.}\label{tab:muscle-parameters}
-  \begin{tabular}{@{}cccccccccc@{}}
-    \toprule
-    \thead{Muscle \\ \quad}
-    & \thead{Maximum Isometric \\ Force (N)}  % math \si not in bold
-    & \thead{Optimal Fiber \\ Length (m)}
-    & \thead{Tendon Slack \\ Length (m)}
-    & \thead{Maximum Contraction \\ Velocity (m / s)} \\
-    \midrule
-    \gls{lr} & 1.4710 & 0.04898 & 0.0084 & 3.8483 \\
-    \gls{mr} & 1.5740 & 0.04084 & 0.0038 & 4.6155 \\
-    \gls{sr} & 1.1768 & 0.04487 & 0.0054 & 4.2009 \\
-    \gls{ir} & 1.4269 & 0.04549 & 0.0048 & 4.1437 \\
-    \gls{so} & 0.6031 & 0.03956 & 0.0265 & 4.7648 \\
-    \gls{io} & 0.5590 & 0.04110 & 0.0015 & 3.5863 \\
-    \bottomrule
-  \end{tabular}
-\end{table}
+created, to avoid abnormal changes on the \gls{fl} curve as the eye rotates.
 
@@
 
-The passive connective tissues apply a restoring force, which brings the eyeball
+The passive connective tissues apply a restoring force, which brings the eye
 back to the central position when the net force from the \gls{em} is zero. These

### 2018-07-18 doc/manuscript.tex
minor updates

@@
 \begin{equation}\label{equ:passive-tissue}
-  \vec{f}_t = -k_p \vec{q} - k_c 10^{-3} \vec{q}^3 - k_v * \vec{\dot{q}}
+  \vec{f}_t = -k_p \vec{q} - k_c \vec{q}^3 - k_d \vec{\dot{q}}
 \end{equation}
@@
 where, $\vec{f}_t$ is the passive tissue forces,
-$k_p= 0.002225 \si{\N \m / \radian}$, $k_c= 34.5297 \si{\N \m / \radian^3}$ and
-$k_v= 0.002$ $\si{\N \m / (\radian / \s)}$ the constants and
-$\vec{\dot{q}} \inr{3}$ the rotational coordinates of the
+$k_p= 2.225 \cdot 10^{-3} \si{\N \m / \radian}$,
+$k_c= 34.53 \cdot 10^{-3} \si{\N \m / \radian^3}$ and
+$k_v= 2 \cdot 10^{-3} \si{\N \m / (\radian / \s)}$ the constants and
+$\vec{q}, \vec{\dot{q}} \inr{3}$ the rotational coordinates of the
 model~\cite{Collins1981}. These forces are modeled using \texttt{OpenSim}'s
@@
 
-A fixation controller was implemented as a custom \texttt{OpenSim} plugin. The
-parameters of the controller are: the desired horizontal $\theta_H$ and vertical
-$\theta_V$ fixation angles (in degrees), the saccade onset and velocity (in
-degrees), and the gains of \gls{pd} tracking controller ($k_p$, $k_d $). A
-sigmoid function is used for generating smooth saccade trajectories
+A fixation controller that calculates the \gls{em} excitations required to
+track a desired saccade was implemented. The controller actuates the model in a
+closed-loop \gls{fd} manner. The parameters of the controller are the desired
+horizontal and vertical fixation angles, the saccade onset and velocity, and the
+gains of \gls{pd} tracking controller. A sigmoid function is used for generating
+smooth saccade trajectories
 
@@
     \theta_d(t) &= \frac{a}{2} \Big(\tanh(b (t - t_0)) + 1\Big) \\
-    \dot{\theta}_d(t) &= \frac{a b}{2} \Big(\tanh^2(b (t - t_0)) - 1\Big)
+    \dot{\theta}_d(t) &= \frac{a b}{2} \Big(1 - \tanh^2(b (t - t_0))\Big)
   \end{aligned}
@@
 $t_0$ a time shift constant. Provided a fixation goal $\theta_g$, a desired
-saccade velocity $\dot{\theta}_g$ and a saccade onset $t_g$ the parameters of
-the sigmoid function are defined as follows $a = \theta_g$,
+saccade velocity $\dot{\theta}_g$ and a saccade onset $t_g$, the parameters of
+the sigmoid function are defined as $a = \theta_g$,
 $b = 2 \dot{\theta}_g / \theta_g$ and $t_0 = t_s$. The \gls{pd} tracking
@@
 
-\begin{equation}\label{equ:pd-controller}
+\begin{equation}\label{equ:eye-pd-controller}
   \ddot{\theta}(t) = k_p (\theta_d(t) - \theta(t)) + k_d (\dot{\theta}_d(t) -
-  \dot{\theta}(t)).
+  \dot{\theta}(t))
 \end{equation}
+%
+where $k_p$, $k_d $ are the tracking gains, and $\theta(t)$, $\dot{\theta}(t)$
+the simulated response of the model.
 
-The sign and magnitude of $\ddot{\theta}(t)$ both for the horizontal and
-vertical displacement of the fixation target is used to actuate the
-corresponding muscles in order to achieve the desired goal. \autoref{fig:model}
-presents an instance of the model during simulation with the corresponding
-muscles activated. \autoref{fig:simulated-saccade} depicts the simulated
-coordinates, speeds and estimated \gls{em} activations that reproduce the
-desired saccade trajectory for different model parameters.
+The sign and magnitude of $\ddot{\theta}(t)$, both for the horizontal and
+vertical displacements, are used to calculate the muscle excitation levels, by
+assuming that each individual muscle rotates the eye globe in a particular
+direction. \autoref{fig:simulated-saccade} depicts the simulated coordinates,
+speeds and estimated \gls{em} excitation levels that reproduce the desired
+saccade trajectory for different model parameters.
 
@@
   \includegraphics[width=.7\textwidth]{eye-model.png}
-  \caption{Model with a fixation at $\theta_H = -15 \si{\degree}$,
-    $\theta_V = 15 \si{\degree}$ during simulation. Blue denotes low and red high
-    muscle activation levels.}\label{fig:model}
+  \caption{Eye model with a fixation $\theta_H = -15 \si{\degree}$,
+    $\theta_V = 15 \si{\degree}$ during simulation. Blue denotes low and red
+    high muscle excitation levels.}\label{fig:model}
 \end{figure}
@@
 \begin{figure}[ht]
-  \vspace{-1.6cm}
+  \vspace{-1.4cm}
   \centering
@@
     keepaspectratio]{UPAT_Eye_Model_Passive_Pulleys_v3_Statesh15v-15kv0002.pdf}\label{fig:simulated-saccade-d}}
-  \caption{Simulated saccade response for different model parameters. The left
-    subplot represents the simulated generalized coordinates; the middle the
-    coordinate speeds; right the estimated \gls{em}
-    activations.}\label{fig:simulated-saccade}
+  \caption{Simulated saccade response for different model parameters. Left
+    subplot represents the simulated generalized coordinates, middle the
+    generalized velocities and right the estimated \gls{em} excitation
+    levels.}\label{fig:simulated-saccade}
 \end{figure}

### 2018-07-18 model/plot_results.py

@@ -90,13 +90,34 @@ def index_containing_substring(list_str, pattern):
     return indices
 
 
+def sigmoid(t, t0, A, B):
+    """Implementation of smooth sigmoid function.
+
+    Parameters
+    ----------
+    t: time to be evalutaed
+    t0: delay
+    A: magnitude
+    B: slope
+
+    Returns
+    -------
+    (y, y', y'')
+
+    """
+    return (A * (np.tanh(B * (t - t0)) + 1) / 2,
+            A * B * (- np.tanh(B * (t - t0)) ** 2 + 1) / 2,
+            - A * B ** 2 * (- np.tanh(B * (t - t0)) ** 2 + 1)
+            * np.tanh(B * (t - t0)))
+
+
 def smooth(x):
     from scipy.signal import medfilt, gauss_spline
-    return medfilt(x, 1)
+    return medfilt(x, 1) # 7
     # return x
 
-###############################################################################
 
+###############################################################################
 
 # state = 'UPAT_Eye_Model_Passive_Pulleys_v3_State_h15v0kv0.sto'
 # state = 'UPAT_Eye_Model_Passive_Pulleys_v3_Statesh0v15kv0.sto'
@@ -111,24 +132,34 @@ coordinates = [1, 2, 3]
 speeds = [7, 8, 9]
 activations = index_containing_substring(labels, 'activation')
 
+theta = np.ceil(np.rad2deg(np.amax(data[:, coordinates])))
+v = np.ceil(np.rad2deg(np.amax(data[:, speeds])))
+t0 = 0.5
+A = theta
+B = 2 * v / theta
+[x, v, a] = sigmoid(time, t0, A, B)
+
 fig, ax = plt.subplots(nrows=1, ncols=3, figsize=(15, 5), sharey=False)
 [
     ax[0].plot(time, np.rad2deg(data[:, coordinate]),
                label=labels[coordinate][2:],)
     for coordinate in coordinates
 ]
+ax[0].plot(time, x, '--', label='desired')
 ax[0].legend()
 ax[0].set_xlabel('time $(s)$')
 ax[0].set_ylabel('coordinates (deg)')
+
 [
     ax[1].plot(time,
                np.rad2deg(data[:, speed]),
                label=labels[speed][2:])
     for speed in speeds
 ]
+ax[1].plot(time, v, '--', label='desired')
[…]

### 2018-07-19 doc/manuscript.tex
filip comments

@@
 \usepackage[pdfpagelabels,pdfusetitle,colorlinks=true,pdfborder={0 0
-  0},pdfauthor={Dimitar Stanev}]{hyperref}
+  0},pdfauthor={Dimitar Stanev},colorlinks=false]{hyperref}
 \usepackage{natbib}
@@
 
-\author{Constantinos Filip, Dimitar Stanev\footnote{Electrical and Computer
+\author{Konstantinos Filip, Dimitar Stanev\footnote{Electrical and Computer
     Engineering Department, University of Patras, Greece; Corresponding
@@
 \begin{abstract}
-  Studying human eye movements has significant implications for improving our
-  understanding of the oculomotor system and treating various visuomotor
-  disorders. An open-source musculoskeletal model of the human eye, that can be
-  used for kinematics and dynamics analysis, is constructed based on the data
-  reported in literature and made publicly available\footnote{SimTK project:
-    \url{https://simtk.org/projects/eye}}. The model is implemented in
+  Physics-based modeling and dynamic simulation of human eye movements has
+  significant implications for improving our understanding of the oculomotor
+  system and treating various visuomotor disorders. We introduce an open-source
+  biomechanical model of the human eye that can be used for kinematics and
+  dynamics analysis. This model is based on the passive pulley hypothesis,
+  constructed based on the data reported in literature regarding physiological
+  measurements of the human eye and made publicly available\footnote{SimTK
+    project: \url{https://simtk.org/projects/eye}}. The model is implemented in
   \texttt{OpenSim}, which is an open-source framework for modeling and
-  simulation of musculoskeletal systems. The calibration of the model parameters
-  is based on physiological measurements of the human eye. The model
-  incorporates an eye globe, orbital suspension tissues and six extraocular
-  muscles. The excitation and activation patterns for a variety of targets can
-  be calculated using the proposed closed-loop fixation controller that drives
-  the model to perform saccadic movements in a forward dynamics manner. The
-  controller minimizes the error between the desired saccadic trajectory and the
-  predicted movement. Consequently, this model enables the investigation muscle
-  activation patterns during static fixation and analyze the dynamics of various
-  eye movements.
+  simulation of musculoskeletal systems. The model incorporates an eye globe,
+  orbital suspension tissues and six extraocular muscles. The excitation and
+  activation patterns for a variety of targets can be calculated using the
+  proposed closed-loop fixation controller that drives the model to perform
+  saccadic movements in a forward dynamics manner. The controller minimizes the
+  error between the desired saccadic trajectory and the predicted
+  movement. Consequently, this model enables the investigation muscle activation
+  patterns during static fixation and analyze the dynamics of eye movements.
 \end{abstract}
@@
 
-Rapid and accurate eye movements are crucial for coordinated direction of
-gaze~\cite{Lee2006}. Studying human eye movement has significant implications
-for improving our understanding of the oculomotor system and treating visuomotor
-disorders. Over the past decades, biomechanics simulation has provided the means
-to analyze different human movements~\cite{Delp2007}. The same principles can be
-used to analyze visual tasks by modeling the musculoskeletal properties of the
+Rapid and accurate eye movements are of great importance for natural vision and
+thus studying human eye movement can improve our understanding of the oculomotor
+system and treating various visuomotor disorders~\cite{Lee2006, Wei2010a}. Over
+the past decades, biomechanics simulation has provided the means to analyze
+different human movements~\cite{Delp2007}. The same principles can be used to
+analyze visual tasks by modeling the musculoskeletal properties of the
 oculomotor system. Consequently, this model can be used to investigate muscle
@@
 movements, calculate metabolic costs and simulate eye disorders, such as
-different forms of strabismus. Furthermore, it can be easily integrated with
-available full body models in order to analyze the relation between the
-vestibular and oculomotor systems.
+different forms of strabismus~\cite{Wong2004}. Furthermore, it can be easily
+integrated with available full body models in order to analyze the relation
+between the vestibular and oculomotor systems.
 
-Saccadic movements are a generated from a coordination of the six
+Eye movements are a generated from the coordinated activation of the six
 \gls{em}. Clinical trials have provided a profound knowledge on the properties
-of the \gls{em} and their line of action on the eye globe~\cite{Robinson1969a},
-the resistive tension of the surrounding tissues~\cite{Collins1981} and the
-length-tension relationship of the muscles~\cite{Iskander2018}. Various
-computational models of the extraocular muscles and orbital mechanics have been
-proposed, which provide insight and scientific bases for oculomotor
-biomechanics, control of eye movement and binocular misalignment. These models
-focus on the realism of muscle behavior and they were based on the viscoelastic
-properties and physiological data \gls{em}.
+of the \gls{em} and their line of action on the eye globe~\cite{Robinson1969a}
+and the resistive tension of the surrounding tissues~\cite{Collins1981,
+  Iskander2018}. Various computational models of the extraocular muscles and
+orbital mechanics have been proposed, which provide insight for oculomotor
+biomechanics, control of eye movement~\cite{Bach-y-Rita1971} and binocular
+misalignment. These models focus on the realism of muscle behavior and they were
+based on the viscoelastic properties and physiological data \gls{em}.
 
@@
 anatomically realistic muscle paths and empirical innervation-length-tension
-relationships. To study the neural control of rapid saccadic movements, models
-using anatomical and mechanical properties of \gls{em} have been developed by
-accounting for the nonlinear muscle dynamics~\cite{Thelen2003a,
+relationships. In order to study the neural control of rapid saccadic movements,
+models using anatomical and mechanical properties of \gls{em} have been
+developed by accounting for the nonlinear muscle dynamics~\cite{Thelen2003a,
   Millard2013}. Such models, having the advantage of supporting dynamics
 simulation, are used in conjunction with brain level
-controllers~\cite{James2018}.
+controllers~\cite{Angelaki2004, James2018}.
 
@@
 controlled by the cranial nerves so as to track a visual target and to stabilize
-the image of the object. The \gls{lr} and \gls{mr} muscles form an antagonistic
-pair that produce horizontal eye movements. The \gls{sr} and \gls{ir} muscles
-form the vertical antagonist pair, which mainly controls vertical eye movement
-and also affects rotation about the horizontal plane and the line of sight
-(secondary action) due to insertion positions. The \gls{so} muscle passes
-through the cartilaginous trochlea attached to the orbital wall, which reflects
-the \gls{so} path by 51 deg. The \gls{io} muscle originates from the orbital
-wall anteroinferior to the globe center and inserts on the sclera posterior to
-the globe equator. The primary actions of \gls{so} and \gls{io} cause rotation
-of the globe around the visual axis and vertical movement.
+the image of the object of interest on the retina. The \gls{lr} and \gls{mr}
+muscles form an agonist/antagonist pair that produce horizontal eye
+movements. The \gls{sr} and \gls{ir} muscles form the vertical
+agonist/antagonist pair, which mainly controls vertical eye movement and also
+affects rotation about the line of sight (secondary action) and the horizontal
+plane (tertiary action). The \gls{so} muscle passes through the cartilaginous
+trochlea attached to the orbital wall, which reflects the \gls{so} path by
+$51 \si{\degree}$. The \gls{io} muscle originates from the orbital wall
+anteroinferior to the globe center and inserts on the sclera posterior to the
+globe equator. The primary actions of \gls{so} and \gls{io} cause rotation of
+the globe around the visual axis, but also affect vertical (secondary action)
+and horizontal (tertiary action) movements.
 
-The model relies on the passive pulley assumption. \autoref{tab:eye-muscle-path}
-shows the positions of the origin, insertion and passive muscle pulleys for the
-\gls{em}, defined in the local body coordinates of the eye globe. The data are
-based on physiological measurements~\cite{Iskander2018}, with some minor
-modification so as to prevent unrealistic muscle-surface penetration. Since no
-position was documented for the origin of the \gls{so}, a point close to the
-origins of the rectus muscles was chosen to match the fiber length in the
-primary position of the \gls{so} muscle.
+The model relies on the passive pulley assumption, which states that the pulleys
+have fixed to the orbit pulley points~\cite{Clark1977,
+  Miller2007}. \autoref{tab:eye-muscle-path} shows the positions of the origin,
+insertion and pulleys for the \gls{em}, defined in the local body coordinates of
+the eye globe. The data are based on physiological
+measurements~\cite{Iskander2018}, with some minor modification so as to prevent
+unrealistic muscle-surface penetration. Since no position was documented for the
+origin of the \gls{so}, a point close to the origins of the rectus muscles was
+chosen to match the fiber length in the primary position of the \gls{so} muscle.
 
@@
     the eye globe (dimensions are given in meters).}\label{tab:eye-muscle-path}
+  \resizebox{\textwidth}{!}{
   \begin{tabular}{@{}cccccccccc@{}}
@@
   \end{tabular}
+  }
 \end{table}
@@
     \gls{so} & 0.6031 & 0.03956 & 0.0265 & 4.7648 \\
-    \gls{io} & 0.5590 & 0.04110 & 0.0015 & 3.5863 \\
+    \gls{io} & 0.5590 & 0.04110 & 0.0015 & 4.5863 \\
     \bottomrule
@@
 
-The \gls{fl} and \gls{fpe} characteristic curves of the \gls{em} differ
+The active \gls{fl} and \gls{fpe} characteristic curves of the \gls{em} differ
 significantly from those of a skeletal muscle. As shown in
@@
 the experimental data available for the \gls{lr} muscle. The values for the
-\gls{fl} and \gls{fpe} characteristic curves are summarized in
-Tables~\ref{tab:eye-fl-curve} and~\ref{tab:eye-fpe-curve}, respectively.
+active \gls{fl} and \gls{fpe} characteristic curves are summarized in
+Tables~\ref{tab:eye-fl-curve} and~\ref{tab:eye-fpe-curve}, respectively. We
+safely assume that the parameters of the characteristic curves for the other
+\gls{em} are the same.
 
@@
     \centering
-    \caption{Parameters of the \gls{fl} characteristic curve for the
+    \caption{Parameters of the active \gls{fl} characteristic curve for the
       \gls{em}.}\label{tab:eye-fl-curve}
@@
   \centering
-  \subfloat[\gls{fl} curve]{\includegraphics[width=0.5\textwidth,
+  \subfloat[active \gls{fl} curve]{\includegraphics[width=0.5\textwidth,
     keepaspectratio]{active-force-length-curve.png}\label{fig:active-force-length-curve}}
@@
     keepaspectratio]{passive-force-length-curve.png}\label{fig:passive-force-length-curve}}
-  \caption{The \gls{fl} and \gls{fpe} curve definition for the Millard muscle
-    model as implemented in \texttt{OpenSim}.}\label{fig:millard-curves}
+  \caption{The active \gls{fl} and \gls{fpe} curve definition for the Millard
+    muscle model as implemented in \texttt{OpenSim}.}\label{fig:millard-curves}
 \end{figure}
@@
 velocity is tuned so as to match the peak velocity of saccadic eye movement
-$\omega^{\text{max}} = 15.7 \si{\radian / \second}$
-($900 \si{\degree / \second}$). Following this definition, the maximum muscle
-contraction velocity is given in optimal fiber length per seconds and it is thus
-different for each \gls{em}, as their optimal fiber length is different
+$\omega^{\text{max}} = 15.7 \si{\radian / \s}$ ($900 \si{\degree /
+  \s}$). Following this definition, the maximum muscle contraction velocity is
+given in optimal fiber length per seconds and it is thus different for each
+\gls{em}, as their optimal fiber length is different
 ($v^{\text{max}} = \omega^{\text{max}} r / l_o^M$). Furthermore, since the optic
 nerve is much shorter that the average muscle nerve, activation and deactivation
-delays ($\tau_a = \tau_d = 5 \si{\milli\second}$) are smaller. Finally, two
+delays ($\tau_a = \tau_d = 5 \si{\milli\s}$) are smaller. Finally, two
 separate wrapping spheres for the rectus muscles and the oblique muscles were
@@
 
-The passive connective tissues apply a restoring force, which brings the eye
-back to the central position when the net force from the \gls{em} is zero. These
-tissues include all non-muscular suspensory tissues, such as Tenon's capsule,
-the optic nerve, the fat pad and the conjunctiva. The force-displacement curve
-of the net elasticity can be represented as
+The passive connective tissues of the orbit apply a restoring force, which
+brings the eye back to the central position when the net force from the \gls{em}
+is zero. These tissues include all non-muscular suspensory tissues, such as
+Tenon's capsule, the optic nerve, the fat pad and the conjunctiva. The
+force-displacement elasticity force can be represented as
 
@@
 $k_c= 34.53 \cdot 10^{-3} \si{\N \m / \radian^3}$ and
-$k_v= 2 \cdot 10^{-3} \si{\N \m / (\radian / \s)}$ the constants and
-$\vec{q}, \vec{\dot{q}} \inr{3}$ the rotational coordinates of the
-model~\cite{Collins1981}. These forces are modeled using \texttt{OpenSim}'s
-expression based coordinate force.
+$k_v= 2 \cdot 10^{-3} \si{\N \m / (\radian / \s)}$ the physiological
+constants~\cite{Robinson1969a, Collins1981, Priamikov2016} and
+$\vec{q}, \vec{\dot{q}} \inr{3}$ the rotational coordinates and velocities of
+the mode. These forces are modeled using \texttt{OpenSim}'s expression based
+coordinate force.
 
@@
 
-A fixation controller that calculates the \gls{em} excitations required to
-track a desired saccade was implemented. The controller actuates the model in a
+A fixation controller that calculates the \gls{em} excitations required to track
+a desired saccade was implemented. The controller actuates the model in a
 closed-loop \gls{fd} manner. The parameters of the controller are the desired
@@
 gains of \gls{pd} tracking controller. A sigmoid function is used for generating
-smooth saccade trajectories
+smooth saccade trajectories in the horizontal and vertical direction, while the
+torsional component is maintained close to zero. More formally,
 
@@
 the sigmoid function are defined as $a = \theta_g$,
-$b = 2 \dot{\theta}_g / \theta_g$ and $t_0 = t_s$. The \gls{pd} tracking
-controller has the following form
+$b = 2 \dot{\theta}_g / \theta_g$ and $t_0 = t_s$. The output of the \gls{pd}
+tracking controller has the following form
 
 \begin{equation}\label{equ:eye-pd-controller}
-  \ddot{\theta}(t) = k_p (\theta_d(t) - \theta(t)) + k_d (\dot{\theta}_d(t) -
+  u(t) = k_p (\theta_d(t) - \theta(t)) + k_d (\dot{\theta}_d(t) -
   \dot{\theta}(t))
@@
 
-The sign and magnitude of $\ddot{\theta}(t)$, both for the horizontal and
-vertical displacements, are used to calculate the muscle excitation levels, by
-assuming that each individual muscle rotates the eye globe in a particular
-direction. \autoref{fig:simulated-saccade} depicts the simulated coordinates,
-speeds and estimated \gls{em} excitation levels that reproduce the desired
-saccade trajectory for different model parameters.
+The sign and magnitude of $u(t)$, representing the deviation from
+the fixation target for each axis of rotation respectively, are used to
+calculate the muscle excitation levels, by assuming that each individual muscle
+rotates the eye globe in a particular direction. \autoref{fig:model} presents an
+instance of the model during simulation with the corresponding muscles
+activated. \autoref{fig:simulated-hor-saccade} depicts the simulated
+coordinates, angular velocities and estimated \gls{em} excitation levels that
+reproduce the desired saccade trajectory for different model
+parameters. Finally, \autoref{fig:simulated-cmplx-saccade} shows alternations in
+the saccadic movements both in the horizontal and vertical direction so as to
+examine the activation and deactivation patterns of the \gls{em}.
 
@@
   \centering
-  \includegraphics[width=.7\textwidth]{eye-model.png}
-  \caption{Eye model with a fixation $\theta_H = -15 \si{\degree}$,
-    $\theta_V = 15 \si{\degree}$ during simulation. Blue denotes low and red
-    high muscle excitation levels.}\label{fig:model}
+  \subfloat{\includegraphics[height=.25\textheight]{model.png}}
+  \hspace{3cm}
+  \subfloat{\includegraphics[height=.25\textheight]{eye-globe.png}}
+  \caption{Model with a fixation target at $\theta_H = -15 \si{\degree}$,
+    $\theta_V = 0 \si{\degree}$ during simulation. Blue denotes low and red high
+    muscle activation levels.}\label{fig:model}
 \end{figure}
@@
 \begin{figure}[ht]
-  \vspace{-1.4cm}
   \centering
-  \subfloat[$\theta_H = 15 \si{\degree}$, $\theta_V = 0 \si{\degree}$, $b = 600
-  \si{\degree / \s}$ and $k_v = 0 \si{\N \m / (\radian / \s)}$]{\includegraphics[width=1.\textwidth,
-    keepaspectratio]{UPAT_Eye_Model_Passive_Pulleys_v3_State_h15v0kv0.pdf}\label{fig:simulated-saccade-a}} 
-  \\
-  \subfloat[$\theta_H = 0 \si{\degree}$, $\theta_V = 15 \si{\degree}$, $b = 600
-  \si{\degree / \s}$ and $k_v = 0 \si{\N \m / (\radian / \s)}$]{\includegraphics[width=1.\textwidth,
-    keepaspectratio]{UPAT_Eye_Model_Passive_Pulleys_v3_Statesh0v15kv0.pdf}\label{fig:simulated-saccade-b}}
-  \\
-  \subfloat[$\theta_H = -15 \si{\degree}$, $\theta_V = 15 \si{\degree}$, $b = 600
-  \si{\degree / \s}$ and $k_v = 0 \si{\N \m / (\radian / \s)}$]{\includegraphics[width=1.\textwidth,
-    keepaspectratio]{UPAT_Eye_Model_Passive_Pulleys_v3_Statesh-15v15kv0.pdf}\label{fig:simulated-saccade-c}}
-  \\
-  \subfloat[$\theta_H = -15 \si{\degree}$, $\theta_V = 15 \si{\degree}$, $b = 100
-  \si{\degree / \s}$ and $k_v = 0.002 \si{\N \m / (\radian / \s)}$]{\includegraphics[width=1.\textwidth,
-    keepaspectratio]{UPAT_Eye_Model_Passive_Pulleys_v3_Statesh15v-15kv0002.pdf}\label{fig:simulated-saccade-d}}
+  \subfloat[$\theta_H = -15 \si{\degree}$, $\theta_V = 0 \si{\degree}$, $b = 600
+  \si{\degree / \s}$ and $k_v = 0.002 \si{\N \m / (\radian / \s)}$]
+  {\includegraphics[trim={3cm 0cm 2.8cm 0cm},clip,width=1.\textwidth]{horizontal_move.pdf}}  \\
+  \subfloat[$\theta_H = -15 \si{\degree}$, $\theta_V = 0 \si{\degree}$, $b = 600
+  \si{\degree / \s}$ and $k_v = 0 \si{\N \m / (\radian / \s)}$]
+  {\includegraphics[trim={3cm 0cm 2.8cm 0cm},clip,width=1.\textwidth]{abd_no_visc.pdf}} 
   \caption{Simulated saccade response for different model parameters. Left
@@
     generalized velocities and right the estimated \gls{em} excitation
-    levels.}\label{fig:simulated-saccade}
+    levels.}\label{fig:simulated-hor-saccade}
 \end{figure}
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[trim={3cm 0cm 2.8cm 0cm},clip,width=1.\textwidth]{complex_move.pdf}
+  \caption{Simulated saccade response performing abduction/adduction
+    ($0-2 \si{\s}$) and supraduction/infraduction ($2-4 \si{\s}$) with
+    $k_v = 0.002 \si{\N \m / (\radian /
+      \s)}$.}\label{fig:simulated-cmplx-saccade}
+\end{figure}
+
+% \begin{figure}[ht]
+%   \centering
+%   \includegraphics[width=.7\textwidth]{eye-model.png}
+%   \caption{Eye model with a fixation $\theta_H = -15 \si{\degree}$,
+%     $\theta_V = 15 \si{\degree}$ during simulation. Blue denotes low and red
+%     high muscle excitation levels.}\label{fig:model}
+% \end{figure}
+
+% \begin{figure}[ht]
+%   \vspace{-1.4cm}
+%   \centering
+%   \subfloat[$\theta_H = 15 \si{\degree}$, $\theta_V = 0 \si{\degree}$, $b = 600
+%   \si{\degree / \s}$ and $k_v = 0 \si{\N \m / (\radian / \s)}$]{\includegraphics[width=1.\textwidth,
+%     keepaspectratio]{UPAT_Eye_Model_Passive_Pulleys_v3_State_h15v0kv0.pdf}\label{fig:simulated-saccade-a}}
+%   \\
+%   \subfloat[$\theta_H = 0 \si{\degree}$, $\theta_V = 15 \si{\degree}$, $b = 600
+%   \si{\degree / \s}$ and $k_v = 0 \si{\N \m / (\radian / \s)}$]{\includegraphics[width=1.\textwidth,
+%     keepaspectratio]{UPAT_Eye_Model_Passive_Pulleys_v3_Statesh0v15kv0.pdf}\label{fig:simulated-saccade-b}}
+%   \\
+%   \subfloat[$\theta_H = -15 \si{\degree}$, $\theta_V = 15 \si{\degree}$, $b = 600
+%   \si{\degree / \s}$ and $k_v = 0 \si{\N \m / (\radian / \s)}$]{\includegraphics[width=1.\textwidth,
+%     keepaspectratio]{UPAT_Eye_Model_Passive_Pulleys_v3_Statesh-15v15kv0.pdf}\label{fig:simulated-saccade-c}}
+%   \\
+%   \subfloat[$\theta_H = -15 \si{\degree}$, $\theta_V = 15 \si{\degree}$, $b = 100
+%   \si{\degree / \s}$ and $k_v = 0.002 \si{\N \m / (\radian / \s)}$]{\includegraphics[width=1.\textwidth,
+%     keepaspectratio]{UPAT_Eye_Model_Passive_Pulleys_v3_Statesh15v-15kv0002.pdf}\label{fig:simulated-saccade-d}}
+%   \caption{Simulated saccade response for different model parameters. Left
+%     subplot represents the simulated generalized coordinates, middle the
+%     generalized velocities and right the estimated \gls{em} excitation
+%     levels.}\label{fig:simulated-saccade}
+% \end{figure}
 
@@
 proposed computational model. In this study, we didn't attempt to model the
-muscle pulleys~\cite{Kono2002a}, which vary as a function of the model
-coordinates. Therefore, the users should consider performing further validation
-of the eye model based on the requirements of the targeted utility and the
-variables of interests.
+muscle pulleys~\cite{Kono2002a}, where the position of the pulleys vary as a
+function of the model coordinates. Therefore, the users should consider
+performing further validation of the eye model based on the requirements of the
+targeted utility and the variables of interests.
 

### 2018-07-19 source_code/src/fixation_controller_plugin/FixationController.cpp

@@ -72,18 +72,35 @@ void FixationController::computeControls(const State& s, Vector& controls) const
                          get_saccade_onset(),
                          thetaH,
                          2 * desSaccadeVelocity / abs(thetaH));
+    // auto sigmY = sigmoid(t,
+    //                      get_saccade_onset(),
+    //                      thetaH,
+    //                      2 * desSaccadeVelocity / abs(thetaH))
+    //   + sigmoid(t - 1,
+    //             get_saccade_onset(),
+    //             -thetaH, 2 * desSaccadeVelocity / abs(thetaH));
 
     double ydes = isnan(sigmY[0]) ? 0 : sigmY[0];
     double ydesv = isnan(sigmY[1]) ? 0 : sigmY[1];
-    double ydesa = sigmY[2];
+    double ydesa = isnan(sigmY[2]) ? 0 : sigmY[2];
 
     auto sigmZ = sigmoid(t,
                          get_saccade_onset(),
                          thetaV,
                          2 * desSaccadeVelocity / abs(thetaV));
+    // auto sigmZ = sigmoid(t,
+    //                      get_saccade_onset(),
+    //                      thetaV - thetaV, 2 * desSaccadeVelocity / abs(thetaV))
+    //   + sigmoid(t - 2,
+    //             get_saccade_onset(),
+    //             thetaV, 2 * desSaccadeVelocity / abs(thetaV))
+    //   + sigmoid(t - 3,
+    //             get_saccade_onset(),
+    //             -thetaV, 2 * desSaccadeVelocity / abs(thetaV));
+
     double zdes = isnan(sigmZ[0]) ? 0 : sigmZ[0];
     double zdesv = isnan(sigmZ[1]) ? 0 : sigmZ[1];
-    double zdesa = sigmZ[2];
+    double zdesa = isnan(sigmZ[2]) ? 0 : sigmZ[2];
 
     // Get the current position and velocity
     double x = xCoord.getValue(s); double xv = xCoord.getSpeedValue(s);

### 2018-07-19 source_code/src/fixation_controller_plugin/FixationController.h

@@ -5,7 +5,7 @@
  * horizontal and vertical target angle and computes the muscle excitation
  * required to track the desired motion.
  *
- * @author Constantinos Filip <filipconstantinos@gmail.com>
+ * @author Konstantinos Filip <filipconstantinos@gmail.com>
  *         Dimitar Stanev     <jimstanev@gmail.com>
  *
  * @see <a href="https://simtk.org/projects/eye">[SimTK Project]</a>, <a

### 2018-07-19 source_code/src/simulation/Simulation.cpp

@@ -6,7 +6,7 @@
  * the model and a forward simulation is carried. All results are stored in the
  * model/ folder.
  *
- * @author Constantinos Filip <filipconstantinos@gmail.com>
+ * @author Konstantinos Filip <filipconstantinos@gmail.com>
  *         Dimitar Stanev     <jimstanev@gmail.com>
  *
  * @see <a href="https://simtk.org/projects/eye">[SimTK Project]</a>, <a

### 2018-07-20 source_code/src/fixation_controller_plugin/FixationController.h
update publication

@@ -9,7 +9,7 @@
  *         Dimitar Stanev     <jimstanev@gmail.com>
  *
  * @see <a href="https://simtk.org/projects/eye">[SimTK Project]</a>, <a
- * href="XXX">[Publication]</a>
+ * href="https://arxiv.org/abs/1807.07332">[Publication]</a>
  */
 #ifndef FIXATION_CONTROLLER_H
 #define FIXATION_CONTROLLER_H

### 2018-07-20 source_code/src/simulation/Simulation.cpp

@@ -10,7 +10,7 @@
  *         Dimitar Stanev     <jimstanev@gmail.com>
  *
  * @see <a href="https://simtk.org/projects/eye">[SimTK Project]</a>, <a
- * href="XXX">[Publication]</a>
+ * href="https://arxiv.org/abs/1807.07332">[Publication]</a>
  */
 #include <iostream>
 #include <OpenSim/OpenSim.h>

### 2018-08-22 model/plot_results.py
fix bug in v4.0

@@ -122,14 +122,23 @@ def smooth(x):
 # state = 'UPAT_Eye_Model_Passive_Pulleys_v3_State_h15v0kv0.sto'
 # state = 'UPAT_Eye_Model_Passive_Pulleys_v3_Statesh0v15kv0.sto'
 # state = 'UPAT_Eye_Model_Passive_Pulleys_v3_Statesh-15v15kv0.sto'
-state = 'UPAT_Eye_Model_Passive_Pulleys_v3_Statesh15v-15kv0002.sto'
+# state = 'UPAT_Eye_Model_Passive_Pulleys_v3_Statesh15v-15kv0002.sto'
+state = 'UPAT_Eye_Model_Passive_Pulleys_v4_States_v4.sto'
 
 header, labels, data = readMotionFile(state)
 
 data = np.array(data)
 time = np.array(data[:, 0])
-coordinates = [1, 2, 3]
-speeds = [7, 8, 9]
+
+# v4 is OpenSim v4.0 compatible thus the state are stored differently from
+# OpenSim v3.3
+if 'v4' in state:  # OpenSim v4.0
+    coordinates = index_containing_substring(labels, 'value')[0:3]
+    speeds = index_containing_substring(labels, 'speed')[0:3]
+else:  # OpenSim v3.3
+    coordinates = [1, 2, 3]
+    speeds = [7, 8, 9]
+
 activations = index_containing_substring(labels, 'activation')
 
 theta = np.ceil(np.rad2deg(np.amax(data[:, coordinates])))
@@ -142,7 +151,7 @@ B = 2 * v / theta
 fig, ax = plt.subplots(nrows=1, ncols=3, figsize=(15, 5), sharey=False)
 [
     ax[0].plot(time, np.rad2deg(data[:, coordinate]),
-               label=labels[coordinate][2:],)
+               label=labels[coordinate])
     for coordinate in coordinates
 ]
 ax[0].plot(time, x, '--', label='desired')
@@ -153,7 +162,7 @@ ax[0].set_ylabel('coordinates (deg)')
 [
     ax[1].plot(time,
                np.rad2deg(data[:, speed]),
-               label=labels[speed][2:])
+               label=labels[speed])
     for speed in speeds
 ]
 ax[1].plot(time, v, '--', label='desired')

### 2018-08-22 source_code_opensim_v4.0/src/fixation_controller_plugin/internal/RegisterPlugin.cpp

@@ -1,6 +1,19 @@
 #include "RegisterPlugin.h"
-#include <OpenSim/Simulation/Model/ExpressionBasedCoordinateForce.h>
+/*
+ * Strange incomplete type error
+ *
+ * I was getting a strange incomplete type error when including
+ * FixationController.h here. The fixation controller includes the Controller.h
+ * file which makes a forward declaration of type Actuator that is passed to the
+ * Set<> as a const Actuator. However, the Set object performs some operations
+ * that violate the constness. Not sure why this triggered an error that is
+ * located in the opensim-core library. However, by including Actuator.h before
+ * the FixationController.h seems to solve the problem. This is not necessary
+ * when using the fixation controller elsewhere (e.g., see Simulation.cpp).
+ */
+#include <OpenSim/Simulation/Model/Actuator.h>
 #include "../FixationController.h"
+#include <OpenSim/Simulation/Model/ExpressionBasedCoordinateForce.h>
 
 using namespace OpenSim;
 
@@ -13,7 +26,7 @@ void RegisterPlugin() {
     // fix OpenSim bug: ExpressionBasedCoordinateForce is not registered, thus
     // it can't be used in the .osim model.
     Object::RegisterType(ExpressionBasedCoordinateForce());
-    // Object::RegisterType(FixationController());
+    Object::RegisterType(FixationController());
 }
 
 dllObjectInstantiator::dllObjectInstantiator() {

### 2018-08-22 source_code_opensim_v4.0/src/include/Settings.h

@@ -3,6 +3,6 @@
 
 #include <string>
 
-std::string MODEL_DIR = "D:/Desktop/upat_eye_model/source_code_opensim_v4.0/../model/";
+std::string MODEL_DIR = "/home/mitkof6/dev/upat_eye_model/source_code_opensim_v4.0/../model/";
 
 #endif
